---
title: "Institutional colors"
author: "SPDS, uni.kn"
date: "2022 12 18"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Institutional colors}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
options(max.print = "75")
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               collapse = TRUE, 
               comment = "#>",
               message = FALSE,
               warning = FALSE,
               ## Default figure options:
               # fig.width = 6, 
               # fig.asp = .8 # .618, # golden ratio
               fig.align = "center",
	             out.width = "60%"
               )
```

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

The color palettes included in the **unikn** package are based on the CD\ manual of the [University of Konstanz](https://www.uni-konstanz.de/). 
However, the functionality provided by the package makes it easy and straightforward to define and use **your own colors**.  

This vignette first shows which color palettes from other institutions are currently provided by the **unikn** package. 
As these will never be complete, it then explains how the color-related tools of **unikn** can be used for **creating new color palettes**. 
<!-- Main color info: --> 
(For a general introduction to the package's colors, color palettes, and color functions, see the vignettes on [Colors and color functions](colors.html) and [Color recipes](color_recipes.Rmd). 
The vignette on [text](text.html) provides information on creating text boxes and decorations.) 

We install or load the **unikn** package to get started: 

```{r load-pkg-colors, message = FALSE, warning = FALSE}
# install.packages('unikn')  # install unikn from CRAN client
library('unikn')             # load the package
```


## Current color palettes

### Overview

In addition to the specific color palettes of the [University of Konstanz](https://www.uni-konstanz.de/), 
**unikn** currently provides the general color palettes of the following institutions: 

- [Caltech](https://www.caltech.edu/): `caltech_pal_1`, `caltech_pal_2`, `caltech_pal_3` 
- [ETH Zurich](https://ethz.ch/): `eth_pal`, `eth_pal_light`, `eth_pal_grey` 
- [Free University Berlin](https://www.fu-berlin.de/en/index.html): `fu_pal_0`, `fu_pal_1`, `fu_pal_2`, `fu_pal_3` 
- [Humboldt University Berlin](https://www.hu-berlin.de/en): `hu_pal_1`, `hu_pal_2` 
- [Max Planck Society](https://www.mpg.de/en): `mpg_pal` 
- [University of Bonn](https://www.uni-bonn.de/en): `uni_bonn_1`, `uni_bonn_2` 
- [University of Freiburg](https://uni-freiburg.de/): `uni_freiburg_br`, `uni_freiburg_blue`, `uni_freiburg_grey`, `uni_freiburg_info` 
- [University of Göttingen](https://uni-goettingen.de/): `uni_goettingen_1`, `uni_goettingen_2`, `uni_goettingen_3` 
- [University of Hamburg](https://www.uni-hamburg.de/): `uni_hamburg_1`, `uni_hamburg_2` 
- [University of Köln](https://www.uni-koeln.de/): `uni_koeln_1`, `uni_koeln_2` 
- [University of Konstanz](https://www.uni-konstanz.de/): `uni_konstanz`, `uni_konstanz_pref` 
- [Princeton University](https://www.princeton.edu/): `uni_princeton_0`, `uni_princeton_1`, `uni_princeton_2` 
- [University of Regensburg](https://www.uni-regensburg.de/): `uni_regensburg_1`, `uni_regensburg_2`, `uni_regensburg_3` 
- [Rensselaer Polytechnic Institute](https://www.rpi.edu/): `rpi_pal_1`, `rpi_pal_2`, `rpi_pal_3`

<!-- Note: Number of palettes & exported as named vectors: -->

These 40\ palettes are exported (as named vectors) and documented (to credit their contributors and sources). 
We can use and modify these palettes with the **unikn** [color functions](colors.html):

<!-- Note: Duplicate image in README.Rmd -->

```{r show-add-pals, fig.width = 7.2, out.width = "550px", fig.asp = 1.5}
seecol("add", main = "Color palettes from various institutions")
```


### Call for contributions

<!-- Collecting contributed color palettes: -->

> **Collecting color palettes**
>
> Are you using **unikn** to create **your own color palettes**?  
> If so, please **let us know** (e.g., by posting on this [GitHub issue](https://github.com/hneth/unikn/issues/11))!  
>  
> We collect color palettes from other institutions to make them available in future versions of **unikn**.  
>   
> To enable us to verify and provide appropriate credit to your contributions, please provide us with the following information:  
>   
> 1. your code (e.g., the `newpal()` command creating your color palette),   
> 2. your source (e.g., some URL with color definitions),   
> 3. your name and some contact information (e.g., an Email address).  
>  
> We're looking forward to your inputs and contributions!  


## Creating new color palettes

A color palette is an ordered collection of color objects. 
Thus, when creating a new color palette, we need to 

1. define its elementary colors as R objects, and then 

2. combine them into a color palette (usually an R data frame or named vector). 

In the following, we illustrate both steps. 
As an example, we will re-create the color palette of the German [Max Planck Society](https://www.mpg.de/en) (that **unikn** exports as `mpg_pal`) from the color definitions in its Corporate Design manual (available [here](https://docplayer.org/2328711-Max-planck-institut-das-erscheinungsbild-der-max-planck-gesellschaft-4-ueberarbeitete-auflage.html)). 


### 1. Defining colors

The representation of color is a wide-ranging topic that has fascinated both scholars and practicioners for centuries. 
Essentially, there are many alternative ways to analyze or synthesize colors, and the most appropriate representational way depends on the specific context of a particular use case (see [Wikipedia: Color model](https://en.wikipedia.org/wiki/Color_model) for an introduction). 

In R, the **grDevices** package (included in any default installation) provides colors and color-related functions. 
Although R supports many color models (including HCL, HSV, Lab, Luv, and RGB variants), the three most common ways of expressing colors in\ R are: 

1. by R color name: 
R includes `r length(colors())` named colors, whose names are provided (as a character vector) by evaluating the `colors()` function of **grDevices**. 

2. by RGB\ values: 
Every color (or hue) is expressed by a triple of numeric values on three dimensions that denote the additive primary colors (red, green, blue). 
The range of values is either expressed on a scale from\ 0 to\ 1, or ---\ more typically\ --- on a scale from\ 0 to\ 255 (i.e., on an 8\ bit scale allowing for $2^8 = 256$ different values). 

3. by HEX\ values: 
Every color (or hue) is expressed by a character string (with the prefix\ `#`), followed by a triple of hexadecimal values (each ranging from\ `00` to\ `FF`) that correspond to the triple of decimal RGB values (in the 0 to 255 range).

<!-- Note: 3 ways to express the same color -->

Importantly, the _same_ color (or hue) can be denoted by its name, its RGB\ values, and its HEX\ values. 
Thus, the three different systems (i.e., color name, RGB values, or HEX values) can be _alternative_ representations of the same object. However, just as the properties of an alphabet or numeral notation system constraints the words or numbers that can be expressed by it, the universe of possible colors is limited by their representation. 
For instance, whereas the RGB and HEX systems usually distinguish between more than 16\ million (i.e., $(2^8)^3 = 256^3 = 16.777.216$) possible colors, only a limited number of them are named. Similarly, when multiple names refer to the same triple of RGB or HEX values, they denote the same color.^[Strictly speaking, different shades of a given color can be further distinguished by different _opacity_ or _transparency_ levels (typically denoted by an\ `alpha` value). However, the visual appearance of a transparent color can be mimicked by mixing the color with the background color (usually "white"). Hence, transparent shades of a color are usually not counted as different colors.]

<!-- Viewing colors (with seecol()):  -->

An example helps clarifying these abstract terms. 
When viewing an existing color or color palette with the `seecol()` function of **unikn**, the name, HEX and RGB values of the individual colors are being shown. 
For instance, the color palette `mpg_pal` provides the colors of the [Max Planck Society](https://www.mpg.de/en) as: 

```{r def-col-mpg-example, fig.width = 5, fig.asp = 1, fig.align = 'center'}
seecol(pal = mpg_pal, col_brd = "black", lwd_brd = .5)
```

As the middle color of `mpg_pal` is one of the `r length(colors())` named colors of `colors()`, R will recognize it as a color when we type its name (as a character object `"white"`) in a context that expects color inputs. 
However, the two other colors of `mpg_pal` have specifically been defined by the [Max Planck Society](https://www.mpg.de/en) and thus are _not_ pre-defined colors of R (i.e., not contained in `colors()`). 
To use such colors in\ R, we first need to define them as color objects. 

<!-- Defining colors in R: -->

So how can we define new colors in R? 
Most institutions will provide definitions of their corporate colors or instructions how they can be defined in some standard color system. 
Unfortunately, some definitions (such as those using the subtractive CMYK model) are aimed at printing (as opposed to computer screens) and other standards (such as numeric PANTONE&trade; PMS codes) are proprietary. 
But as most institutions have an interest in widely disseminating their colors, they typically include translations or approximations for alternative color models in their definitions. 

<a href="https://www.mpg.de/en">
<img src = "../inst/pix/logo_MPG.png" align = "right" alt = "MPG" width = "150px" style = "border:10;"/>
</a>

The [Max Planck Society](https://www.mpg.de/en) defines two main colors in its CD\ manual (available [here](https://docplayer.org/2328711-Max-planck-institut-das-erscheinungsbild-der-max-planck-gesellschaft-4-ueberarbeitete-auflage.html)): 

- A dark shade of green is defined as Pantone\ 328, and approximated by CMYK\ 100/0/57/30 or RGB\ 17/102/86. 

- A light shade of grey is Pantone\ 427, and approximated by CMYK\ 0/0/6/15 or RGB\ 221/222/214. 

In the following, we will first use the RGB values to define both colors (as these are provided in the definition).
However, we then show that we alternatively could have used the colors' HEX\ values (had they been provided) to define the same color objects.

#### Defining colors from their RGB values {-}

As our two colors of interest are given in RGB values, we can use the `rgb()` function of the **grDevices** package to define corresponding objects:

```{r def-col-rgb}
mpg_green <- grDevices::rgb( 17, 102,  86, maxColorValue = 255)
mpg_grey  <- grDevices::rgb(221, 222, 214, maxColorValue = 255)
```

The first three arguments specify values on the `red`, `green`, and `blue` dimension, respectively. 
Importantly, the values must be in the range\ $[0, max]$, where $max$ is provided by `maxColorValue = 255`. 
Whereas the [sRGB standard](https://en.wikipedia.org/wiki/SRGB) expresses all values in a normalized range of\ $[0, 1]$, most color definitions use the traditional 8-bit representation that distinguish between $2^8 = 256$\ levels in the range\ $[0, 255]$. 
The `rgb()` function additionally allows for adding\ `names` and a transparency level\ `alpha`, but as we can modify these aspects later, we do not need to use them at this point. 

What do we obtain when we evaluate our new color objects? 
Let's see: 

```{r def-col-rgb-eval}
mpg_green
mpg_grey
```

When evaluating an object created by `rgb()`, R returns a character sequence of the form\ `"#rrggbb"`. 
In this string, the sequence\ `rrggbb` expresses the three RGB values (for the `red`, `green`, and `blue` dimension, respectively) in hexadecimal (HEX) notation (in the range $[0, 255]$). 
In HEX notation, the ten numeral symbols (i.e., the digits from\ 0 to\ 9) or the decimal system are extended by the first six letters of the alphabet (i.e., A to\ F) to turn our usual base-10 notation into an analogous place-value notation with a base of\ 16. 
Thus, the HEX value of `"#116656"` is merely a more compact way of expressing 

- `red`: $1 \cdot 16 + 1 = 17$
- `green`: $6 \cdot 16 + 6 = 102$
- `blue`: $5 \cdot 16 + 6 = 86$


#### Defining colors from their HEX values {-}

The equivalence of the RGB and HEX notations can be illustrated by defining the same colors a second time, but this time specifying their HEX values (returned above):

```{r def-col-hex}
mpg_green_2 <- "#116656"
mpg_grey_2  <- "#DDDED6"
```

If both definitions really yield identical R objects, the following checks must evaluate to `TRUE`:

```{r def-col-rgb-hex-eual}
all.equal(mpg_green, mpg_green_2)
all.equal(mpg_grey,  mpg_grey_2)
```

Thus, we can define colors from both their RGB or HEX values ---\ and the two notations are really just two alternative ways of representing the same colors. 


### 2. Combining colors into a palette 

Having defined two new colors, we can easily combine them into a new color palette. 
A color palette must be stored in some sequential data structure. 
The two simplest data structures in\ R that are suited for this purpose are (named) vectors or data frames. 
Thus, the following two expressions both define valid color palettes: 

```{r cols-to-pals}
c(mpg_green, mpg_grey)
data.frame(mpg_green, mpg_grey)
```

+++ here now +++ 

**ToDo**:

- Using `newpal()` simplifies creating named palettes (as data frames or named vectors)
- Mixing colors and creating color gradients with `usecol()`. 


#### Notes {-}

Getting 50% shades:

```{r def-col-rgb-50pc}
mpg_green_50 <- grDevices::rgb( 17, 102,  86, alpha = 256 * .50, names = "MPG green (alpha = 50%)", maxColorValue = 255)
```

Approximation: Create gradients of non-transparent colors by mixing with `"white"`

```{r def-mix-with-white}
# Create gradient of 3 colors (from mpg_green to "white") and select middle one:
mpg_green_50_mix <- usecol(c(mpg_green, "white"), n = 3)[2]
```

Comparing both results reveals differences:

```{r def-compare, fig.width = 5, fig.asp = .90, fig.align = 'center'}
seecol(c(mpg_green_50, mpg_green_50_mix))
```


Note on finding custom color names:

```{r eval = FALSE}
# Note: Names of use-defined colors are not found in environment, but: 
all_colors()[names(all_colors()) == "MPG green"]
```


### Examples

Representing basic colors (by RGB and HEX values):

A. Maximize contrast: black vs. white: 

```{r rgb-col-def-A, eval = FALSE}
# A: Black vs. white: Maximize contrast (on all 3 dimensions): ------ 

c_000 <- rgb(red = 0, green = 0, blue = 0, alpha = NULL, names = "Color 000", maxColorValue = 1)
c_111 <- rgb(red = 1, green = 1, blue = 1, alpha = NULL, names = "Color 111", maxColorValue = 1)

# View colors:
unikn::seecol(c(c_000, c_111), 
              col_brd = "grey33", lwd_brd = .50,
              main = "Extreme RGB values")

# Note intermediate cases: Shades of grey
seecol(usecol(c(c_000, c_111), n = 5), 
              col_brd = "grey33", lwd_brd = .50,
              main = "Shades of grey")
```

B. Extreme RGB values (on 1 dimension): 

```{r rgb-col-def-B, eval = FALSE}
# B: Extreme RGB values (on 1 dimension): ---- 

c_100 <- rgb(red = 1, green = 0, blue = 0, alpha = NULL, names = "Color 100", maxColorValue = 1)
c_010 <- rgb(red = 0, green = 1, blue = 0, alpha = NULL, names = "Color 010", maxColorValue = 1)
c_001 <- rgb(red = 0, green = 0, blue = 1, alpha = NULL, names = "Color 001", maxColorValue = 1)

# View extremes: 
seecol(c(c_100, c_010, c_001), 
       # col_brd = "grey33", lwd_brd = .50,
       main = "Extremes on 1 dimension")
```

Combining A. and B.:

```{r rgb-col-def-C, eval = FALSE}
# A and B:
seecol(c(c_000, c_100, c_010, c_001, c_111), 
       col_brd = "grey33", lwd_brd = .50,
       main = "5 corners of the RGB color cube")
```

We see:

- Standard range from 0 to 255 (8-bit)
- Same values on all dimensions: 256 shades of grey
- An additive model: Complete absence of colors is "black", maximum presence of all colors is "white".
- 3 complements of corners are CMY (subtractive color model)


#### Explaining HEX values {-}

The decimal notation system uses different symbols (aka. digits or numerals) to represent 10\ distinct numeric values from\ $0$ to\ $9$ and a place-value notation to express numeric values beyond\ $9$ as abbreviated polynomials. For instance, the three digit decimal notation\ $123$ is actually an abbreviation for the numeric value of $1 \cdot 10^2 + 2 \cdot 10^1 + 3 \cdot 10^0$. 

A hexadecimal notation system (HEX) uses the same 10 initial symbols, but extends their range by using different symbols for 16 distinct numeric values from\ $0$ to\ $15$. The decimal values\ $10$ to\ $15$ are represented by the letters\ $A$ to\ $F$. 
Numeric values of 16 and beyond are expressed by using the same polynomial convention. 
Thus, the two digit sequence\ $10$ (in HEX notation) now represents a numeric value of\ 16 (in decimal notation), and the numeric value of the digit sequence\ $123$ (in HEX notation) represents a value of $1 \cdot 16^2 + 2 \cdot 16^1 + 3 \cdot 16^0 = 291$ (in decimal notation). 

The following Table lists the first 20\ symbols (or symbol combinations, for numeric values ranging from zero to nineteen) in both the decimal and HEX notation: 

<!-- Table: -->

Symbol Nr.:       |  1  |  2  |  3  |  4  |  5  |   6  | 7  |  8  |  9  | 10  | 11  | 12  | 13  | 14  | 15  | 16  | 17  | 18  | 19  | 20  | 
:---------------  | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | --: | 
Value in decimal notation: |  0 |  1 |  2 |  3 |  4 |  5 |  6 |  7 |  8 |  9 |  10 |  11  | 12  | 13  | 14  | 15  | 16  |  17 |  18 |  19 | 
Value in HEX notation:     |  0 |  1 |  2 |  3 |  4 |  5 |  6 |  7 |  8 |  9 |   A |   B  |  C  |  D  |  E  |  F  | 10  |  11 |  12 |  13 | 



```{r dec-hex-base-conversions, eval = FALSE}
library(ds4psy)

# Decimal to hexadecimal (base 16):
dec2base( 9, base = 16)
dec2base(10, base = 16)
dec2base(15, base = 16)
dec2base(16, base = 16)
dec2base(20, base = 16)

# Base 16 to decimal:
base2dec("10", base = 16)  # 16 x  1
base2dec("F0", base = 16)  # 16 * 15
base2dec("F9", base = 16)  # 16 * 15 +  9
base2dec("FF", base = 16)  # 16 * 15 + 15
base2dec("123", base = 16) # (1 * 16^2) + (2 * 16^1) + (3 * 16^0) 
```

Subtle shades of "black" or "white" (in HEX and RGB values):

```{r color-def, echo = FALSE, fig.width = 5, fig.asp = .80, fig.align = 'center'}
# Range of possible colors:
2^8    # 8-binary digits per dimension: 256 values
256^3  # => over 16 million different colors

# Resolution of HEX and RGB:
seecol(c("black", "#010101", "#020202", "grey1", "#040404", "grey2", "#060606", "#070707", "grey3"), 
       col_brd = "white", lwd_brd = .5, main = "A. Black end of 'grey' range")
seecol(c("white", "#FEFEFE", "#FDFDFD", "grey99", "#FBFBFB", "grey98", "#F9F9F9", "#F8F8F8", "grey97"), 
       col_brd = "black", lwd_brd = .5, main = "B. White end of 'grey' range")
```



## Defining color palettes 

The following examples show how to define color palettes (with dedicated color names) for three institutions. 
They can be easily adapted to any other group, company, or entity, that uses some pre-defined set of colors. 

```{r pal-cardiff, echo = FALSE, eval = FALSE, fig.align = 'center', fig.width = 5, fig.height = 3}
# Cardiff University, UK: 
# ToDo. 
```

### A. [University of Freiburg](https://uni-freiburg.de/), Germany: 
<a href="https://uni-freiburg.de/">
<img src = "../inst/pix/logo_freiburg.png" align = "right" width = "120px" alt = "Uni Freiburg" style = "border:10;"/> 
</a>

The [Albert-Ludwigs Universität Freiburg](https://uni-freiburg.de/) provides fairly extensive information on its corporate color scheme (available at `portal.uni-freiburg.de/cmsforum/wsg/webstyleguide/farben` on 2022-10-20). 
Colors can be defined in a variety of ways, but R comes with convenient tools (like the `col2rgb()` and `convertColor()` functions of **grDevices**) to handle most cases. 
The most straightforward way of creating a new color palette in R is by using its HEX/HTML code (provided in character format). 
As the [University of Freiburg](https://uni-freiburg.de/) kindly provides their colors in this format, we can easily define the corresponding color palettes as named vectors: 

```{r freiburg-cols, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
# Source: http://portal.uni-freiburg.de/cmsforum/wsg/webstyleguide/farben

# Basic colors: ---- 

uni_freiburg_br <- c("#004a99", "#c1002a")
names(uni_freiburg_br) <- c("uni-blau", "uni-rot")

uni_freiburg_bwr <- c("#004a99", "white", "#c1002a")  # add "white" for better gradients
names(uni_freiburg_bwr) <- c("uni-blau", "weiss", "uni-rot")

# Web colors: ----

uni_freiburg_blue <- c("#004a99", "#2a6ebb", "#6f9ad3")
names(uni_freiburg_blue) <- c("blue-1", "blue-2", "blue-3")

uni_freiburg_grey <- c("#f2f3f1", "#e0e1dd", "#d5d6d2", "#c9cac8", 
                       "#b2b4b3", "#9a9b9c", "#747678", "#363534")
names(uni_freiburg_grey) <- c("grey-0", "grey-1", "grey-2", "grey-3", 
                              "grey-5", "grey-7", "grey-9", "grey-font")

# Visualizations / illustrations: ----

uni_freiburg_info <- c("#2a6ebb", "#a7c1e3", "#7b2927", "#de3831", "#739600", "#92d400", 
                       "#4d4f53", "#747678", "#b2b4b3", "#d5d6d2", "#e98300", "#efbd47")
names(uni_freiburg_info) <- c("mid-blau", "hell-blau", "dark-red", "hell-red", "mid-green", "hell-green", 
                              "anthrazit", "dark-grey", "mid-grey", "hell-grey", "orange", "gelb")
```

Alternatively, we can define both colors and names in one step by using the `newpal()` function:

```{r freiburg-newpal}
uni_freiburg_info <- newpal(col = c("#2a6ebb", "#a7c1e3", "#7b2927", "#de3831", "#739600", "#92d400", 
                                    "#4d4f53", "#747678", "#b2b4b3", "#d5d6d2", "#e98300", "#efbd47"),
                            names = c("mid-blau", "hell-blau", "dark-red", "hell-red", "mid-green", "hell-green", 
                                      "anthrazit", "dark-grey", "mid-grey", "hell-grey", "orange", "gelb")
                            )
```

Using `newpal()` makes it easy to combine colors and their names and return the resulting color palette in either data frame or vector format.

Once a new color palette has been defined (in your current R\ environment), use all color-related functions to view, modify, and use the palette: 

```{r freiburg-seecol, fig.width = 5, fig.asp = .80, fig.align = 'center', collapse = TRUE}
# View color palette:
seecol(uni_freiburg_info, main = "Colors of the University of Freiburg")
# seecol(uni_freiburg_bwr, n = 7)  # extend color palette
# seecol(c(uni_freiburg_blue, "white", uni_freiburg_grey))  # mix color palette

# use in a demo plot:
demopal(uni_freiburg_info,
        type = 4, col_par = NA, seed = 79098, 
        main = "Colors of the University of Freiburg")
```


### B. [Princeton University](https://princeton.edu/), USA: 
<a href="https://princeton.edu/">
<img src = "../inst/pix/logo_princeton.jpg" align = "right" width = "180px" alt = "Princeton University" style = "border:10;"/>
</a>

The color scheme of [Princeton University](https://princeton.edu/) is easily recognized by its combination of orange with black and white elements. 
The official guidelines (available [here](https://communications.princeton.edu/guides-tools/logo-graphic-identity)) define "Princeton Orange" as Pantone (PMS) 158\ C. 

- The PANTONE&trade; color finder at <https://www.pantone.com/connect/158-C> yields the following color values:

    - RGB: `232 119 34`  
    - HEX/HTML: `#E87722`  
    - CMYK: `0 62 95 0`  

- However, the guide also specifies and distinguishes between two additional versions of orange (depending on background color) and provides the following HEX/HTML values for them:  

    - Orange on white: "#E77500"  
    - Orange on black: "#F58025"  

These definitions suggest defining three separate versions of orange and corresponding color palettes: 

```{r princeton-1, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
# 3 shades of orange: 
orange_basic <- "#E87722"  # Pantone 158 C
orange_white <- "#E77500"  # orange on white background
orange_black <- "#F58025"  # orange on black background

# Defining color palettes:
uni_princeton <- c(orange_basic, "black")
names(uni_princeton) <- c("orange", "black")

# Define color palette and names (in 2 steps):
uni_princeton_1 <- c(orange_white, "white", "black")
names(uni_princeton_1) <- c("orange_w", "white", "black")
```

Again, we can define a palette of named colors in one step:

```{r princeton-2, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
# Basic colors:
uni_princeton_0 <- newpal(col = c(orange_basic, "black"),
                          names = c("orange", "black"))

# 1st color palette (on white backgrounds):
uni_princeton_1 <- newpal(col = c(orange_white, "white", "black"),
                          names = c("orange_w", "white", "black"))
                          
# 2nd color palette (on black backgrounds):
uni_princeton_2 <- newpal(col = c(orange_black, "black", "white"), 
                          names = c("orange_b", "black", "white"))
```

and then inspect a new color palette (with a custom background color and title):

```{r princeton-3, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
seecol(uni_princeton_1, col_bg = "grey90", 
       main = "Colors of Princeton University")
```


### C. [Max Planck Society](https://www.mpg.de/en), Germany: 
<a href="https://www.mpg.de/en">
<img src = "../inst/pix/logo_MPG.png" align = "right" alt = "MPG" width = "150px" style = "border:10;"/>
</a>

The CD manual (available [here](https://docplayer.org/2328711-Max-planck-institut-das-erscheinungsbild-der-max-planck-gesellschaft-4-ueberarbeitete-auflage.html)) of the [Max Planck Society](https://www.mpg.de/en) specifies the use of two primary colors: 

- Green as Pantone 328: Using the PANTONE&trade; color finder at <https://www.pantone.com/connect/328-C> yields the following color values:

    - RGB: `0 115 103`    
    - HEX/HTML: `#007367`   
    - CMYK: `100 10 61 38`  

- Grey as Pantone 427: Using the PANTONE&trade; color finder at <https://www.pantone.com/connect/427-C> yields the following color values: 

    - RGB: `208 211 212`  
    - HEX/HTML: `#D0D3D4`   
    - CMYK: `7 3 5 8`   

```{r mpg-rgb-cols, echo = FALSE, eval = FALSE, fig.width = 6, fig.asp = .618, fig.align = 'center', collapse = TRUE}
# Using RGB values: 
mpg_pal <- data.frame(# named colors: 
  "mpg_green" = rgb( 17, 102,  86, maxColorValue = 255),
  "mpg_gray"  = rgb(221, 222, 214, maxColorValue = 255),
  "mpg_grey"  = rgb(208, 211, 212, maxColorValue = 255)
  )
seecol(pal = list(mpg_pal))  # Ooops...
```

```{r mpg-hex-cols, echo = FALSE, eval = FALSE}
# Using HEX values:
mpg_green <- "#007367"
mpg_grey  <- "#D0D3D4"

# col2rgb(mpg_green)
# col2rgb(mpg_grey)

mpg_pal <- usecol(pal = c(mpg_green, "white", mpg_grey))
# seecol(mpg_pal)
```

Again, the easiest way of defining a corresponding color palette is by defining the elementary colors and then 
combine them into a named color palette by using the `newpal()` function. 
To enable color gradients, we insert the color\ `"white"` between the two\ dedicated colors: 

```{r mpg-newpal}
# Main colors (using HEX values):
mpg_green <- "#007367"
mpg_grey  <- "#D0D3D4"

# Create new color palette (as a vector):
mpg_pal <- newpal(col = c(mpg_green, "white", mpg_grey), 
                  names = c("mpg green", "white", "mpg grey")
                  )
```

As before, can now use the `seecol()` and `usecol()` functions to view, modify, and use the new `mpg_pal` color palette: 

```{r mpg-seecol, fig.width = 5, fig.asp = .80, fig.align = 'center'}
seecol(mpg_pal,
       col_brd = "black", lwd_brd = 1, 
       main = "Colors of the Max Planck Society")
```


## Comparing custom color palettes 

Once a new palette has been defined (and is available in your current R environment), the `seecol()` and `usecol()` functions allow comparing, scaling, and using them in the same fashion as the native **unikn** color palettes: 

```{r compare-custom-pals, eval = TRUE, fig.width = 5, fig.asp = .50, fig.align = 'center', collapse = TRUE}
# Compare and extend custom color palettes:
# seecol(list(uni_freiburg_info, uni_freiburg_bwr, mpg_pal, uni_princeton_1))

# Scaled version: 
seecol(list(uni_freiburg_bwr, uni_princeton_1, mpg_pal, pal_unikn), n = 9,
       pal_names = c("Uni Freiburg", "Princeton Uni", "Max Planck", "Uni Konstanz"),
       main = "Comparing scaled color palettes")
```


## Resources

The following versions of **unikn** and corresponding resources are currently available: 

Type:                    | Version:           | URL:                           |         
:------------------------|:-------------------|:-------------------------------| 
A. **unikn** (R package): | [Release version](https://CRAN.R-project.org/package=unikn) | <https://CRAN.R-project.org/package=unikn> |
    &nbsp;                | [Development version](https://github.com/hneth/unikn/)      | <https://github.com/hneth/unikn/> | 
B. Online documentation:  | [Release version](https://hneth.github.io/unikn/)           | <https://hneth.github.io/unikn/> | 
    &nbsp;                | [Development version](https://hneth.github.io/unikn/dev/)   | <https://hneth.github.io/unikn/dev/> | 


## Vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a> 

The following vignettes provide instructions and examples for using the **unikn** colors, color palettes, and functions: 

<!-- Table of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color functions |
| 2. | [Color recipes](color_recipes.html) | Recipes for color-related tasks |
| 3. | [Institutional colors](color_inst.html) | Creating color palettes for other institutions |
| 4. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->