---
title: "Colors and color palettes"
author: "SPDS, uni.kn"
date: "2019 03 21"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(max.print = "75")
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               collapse = TRUE, 
               comment = "#>",
               message = FALSE,
               warning = FALSE,
               ## Default figure options:
               # fig.width = 7.2, 
               # fig.asp = .618, # golden ratio
               # out.width = "75%",
               fig.align = "center"
               )
```

Explain which colors and color palettes exist in `unikn` and how to use them.

Please load the `unikn` package to get started: 

```{r load_pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```

## Viewing color palettes

Viewing all available color palettes:

```{r seecol_all, fig.width = 7, fig.height = 6}
seecol()  # same as seecol("all")
```


Viewing subsets of the available color palettes: 

Basic palettes:

```{r seecol_unikn_all, fig.width = 7, fig.height = 6}
seecol("unikn_all")
```

Existing gradients:

```{r seecol_grad_all, fig.width = 7, fig.height = 6}
seecol("grad_all")
```



Viewing the colors and details of a particular color palette:

```{r view_pal, fig.width = 7, fig.asp = .618}
seecol(pal_unikn_plus)
```

This cannot only be done by palette object but also by 


* incomplete object name:  __[SHOULD WE EVEN DOCUMENT THE INCOMPLETE CASES?]__

```{r}
seecol(bordeaux)
```


* name: 

```{r}
seecol("pal_bordeaux")
```

* incomplete name: 

```{r}
seecol("bordeaux")
```

## Using colors and color palettes

Using colors in a graph:

```{r use_col, fig.width = 7, fig.asp = .618}
# Using a color palette:
barplot(1/sqrt(1:11), col = seecol(pal_unikn_plus))
```

## Modifying color palettes

Using only a subset of colors:

```{r use_subset, fig.width = 7, fig.height = 6}
seecol(n = 4)
seecol(pal_unikn_plus, 4)
```

Extending a palette to a larger number of colors than contained in the palette:

```{r use_superset, fig.width = 7, fig.height = 6}
seecol(pal = "all", n = 12)
seecol(pal_seeblau, 12)
```

## Mixing color palettes

Combining colors and color palettes: 

```{r combine_pals, fig.width = 7, fig.asp = .618}
seecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
seecol(pal = c(rev(pal_petrol), "white", pal_bordeaux))
```

Combining colors and color palettes to get a specific number of colors: 

```{r combine_pals_n, fig.width = 7, fig.asp = .618}
seecol(pal = col_scale(c(Karpfenblau, Seeblau, "gold"), n = 10))
```

Examples of combining colors and color palettes into new color palettes:

```{r bar_seeblau_pinky, fig.width = 7, fig.asp = .618}
barplot((-5:5)^2+5, col = seecol(pal = c(rev(pal_seeblau), "white", pal_pinky)))
barplot(1/sqrt(11:1), col = seecol(pal = c(rev(pal_seeblau), "white", pal_seegruen)))
```

## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color palettes | 
| 2. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->
