---
title: "Institutional colors"
author: "SPDS, uni.kn"
date: "2023 07 26"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Institutional colors}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
options(max.print = "75")

knitr::opts_chunk$set(echo = TRUE,
                      cache = FALSE,
                      prompt = FALSE,
                      tidy = FALSE,
                      collapse = TRUE, 
                      comment = "#>",
                      message = FALSE,
                      warning = FALSE,
                      # Default figure options:
                      # fig.width = 6, 
                      # fig.asp = .80 # .618, # golden ratio
                      fig.align = "center",
                      out.width = "60%"
)
```

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

The color palettes included in the **unikn** package are based on the CD\ manual of the [University of Konstanz](https://www.uni-konstanz.de/). 
However, the functionality provided by the package makes it easy and straightforward to define and use **your own colors**.  

<!-- Mention the **unicol** package here! -->

This vignette first shows which color palettes from other institutions are currently provided by the **unikn** package. 
As these will never be complete, it then explains how the color-related tools of **unikn** can be used for **creating new color palettes**. 
<!-- Main color info: --> 
(For a general introduction to the package's colors, color palettes, and color functions, see the vignettes on [Colors and color functions](colors.html) and [Color recipes](color_recipes.html). 
The vignette on [Text boxes and decorations](text.html) provides information on plotting slides and text elements.) 

We install or load the **unikn** package to get started: 

```{r load-pkg-colors, message = FALSE, warning = FALSE}
# install.packages('unikn')  # install unikn from CRAN client
library('unikn')             # load the package
```


## Current color palettes

### Overview

In addition to the specific color palettes of the [University of Konstanz](https://www.uni-konstanz.de/), 
**unikn** currently provides the general color palettes of the following institutions: 

- [Caltech](https://www.caltech.edu/): `caltech_pal_1`, `caltech_pal_2`, `caltech_pal_3` 
- [ETH Zurich](https://ethz.ch/): `eth_pal_1`, `eth_pal_2`, `eth_pal_3` 
- [Free University Berlin](https://www.fu-berlin.de/en/index.html): `fu_pal_0`, `fu_pal_1`, `fu_pal_2`, `fu_pal_3` 
- [Humboldt University Berlin](https://www.hu-berlin.de/en): `hu_pal_1`, `hu_pal_2` 
- [Max Planck Society](https://www.mpg.de/en): `mpg_pal` 
- [LMU Munich](https://www.lmu.de/en/index.html): `lmu_pal_1`, `lmu_pal_2`, `lmu_pal_3` 
- [University of Bonn](https://www.uni-bonn.de/en): `uni_bonn_1`, `uni_bonn_2` 
- [University of Freiburg](https://uni-freiburg.de/): `uni_freiburg_0`, `uni_freiburg_1`, `uni_freiburg_2`,   
  (from\ 2022: `uni_freiburg_br`, `uni_freiburg_blue`, `uni_freiburg_grey`, `uni_freiburg_info`) 
- [University of Göttingen](https://uni-goettingen.de/): `uni_goettingen_1`, `uni_goettingen_2`, `uni_goettingen_3` 
- [University of Hamburg](https://www.uni-hamburg.de/): `uni_hamburg_1`, `uni_hamburg_2` 
- [University of Jena](https://www.uni-jena.de/): `uni_jena_1`, `uni_jena_2` 
- [Kiel University](https://www.uni-kiel.de/en/): `uni_kiel_1`, `uni_kiel_2` 
- [University of Köln](https://www.uni-koeln.de/): `uni_koeln_1`, `uni_koeln_2` 
- [University of Konstanz](https://www.uni-konstanz.de/): `uni_konstanz_1`, `uni_konstanz_2` 
- [University of Mannheim](https://www.uni-mannheim.de/): `uni_mannheim_1`, `uni_mannheim_2` 
- [Princeton University](https://www.princeton.edu/): `uni_princeton_0`, `uni_princeton_1`, `uni_princeton_2` 
- [University of Regensburg](https://www.uni-regensburg.de/): `uni_regensburg_1`, `uni_regensburg_2`, `uni_regensburg_3` 
- [University of Ulm](https://www.uni-ulm.de/): `uni_ulm_1`, `uni_ulm_2` 
- [Rensselaer Polytechnic Institute](https://www.rpi.edu/): `rpi_pal_1`, `rpi_pal_2`, `rpi_pal_3` 
- [RPTU Kaiserslautern-Landau](https://rptu.de/): `rptu_pal` 

<!-- Note: Number of palettes & exported as named vectors: -->

These 52\ palettes are exported (as named vectors) and documented (to credit their contributors and sources). 
We can use and modify these palettes with the **unikn** [color functions](colors.html):

<!-- Note: Duplicate image in README.Rmd -->

```{r show-add-pals, fig.width = 7.2, out.width = "550px", fig.asp = 2.0}
seecol("add", main = "Color palettes from various institutions")
```


### Call for contributions

<!-- Collecting contributed color palettes: -->

> **Collecting color palettes**
>
> Are you using **unikn** to create **your own color palettes**?  
> If so, please **let us know** (e.g., by posting on this [GitHub issue](https://github.com/hneth/unikn/issues/11))!  
>  
> We collect color palettes from other institutions to make them available in future versions of **unikn**.  
>   
> To enable us to verify and provide appropriate credit to your contributions, please provide us with the following information:  
>   
> 1. your code (e.g., the `newpal()` command creating your color palette),   
> 2. your source (e.g., some URL with color definitions),   
> 3. your name and some contact information (e.g., an Email address).  
>  
> We're looking forward to your inputs and contributions!  


## Creating new color palettes

A color palette is an ordered collection of color objects. 
Creating new color palettes in R is simple and only requires two steps: 

1. Define the palette's elementary colors as R objects. 

2. Combine colors into a color palette. 

In the following, we illustrate both steps. 
As we will see, Step\ 1 can either be omitted (when the to-be-combined colors exist as R objects) or can be included in Step\ 2.

As an example, we will re-create the color palette of the German [Max Planck Society](https://www.mpg.de/en) (that **unikn** exports as `mpg_pal`) from the color definitions in its Corporate Design manual (available [here](https://docplayer.org/2328711-Max-planck-institut-das-erscheinungsbild-der-max-planck-gesellschaft-4-ueberarbeitete-auflage.html)). 


### 1. Defining colors

Before we can a create new color palette in\ R, its elementary colors need to exist as R\ objects. 
(If all colors to-be-combined into a new palette exist as R\ objects, we can skip this step and directly proceed to **2.\ Combining colors into a palette**.)
Defining new colors in\ R requires a basic understanding of their underlying representation. 

How colors are represented and perceived is a wide-ranging topic that has fascinated scholars and practitioners for centuries. 
Essentially, there are many alternative ways to analyze or synthesize colors, and the most appropriate representational way depends on the specific context of a particular use case (see [Wikipedia: Color model](https://en.wikipedia.org/wiki/Color_model) for an introduction). 

In R, the **grDevices** package (included in any default installation) provides colors and color-related functions. 
Although R supports many color models (including HCL, HSV, Lab, Luv, and RGB variants), the three most common ways of expressing colors in\ R are: 

1. by R color name: 
R includes `r length(colors())` named colors, whose names are provided (as a character vector) by evaluating the `colors()` function of **grDevices**. 

2. by RGB\ values: 
Every color (or hue) is expressed by a triple of numeric values on three dimensions that denote the additive primary colors (red, green, blue). 
The range of values is either expressed on a scale from\ 0 to\ 1, or ---\ more typically\ --- on a scale from\ 0 to\ 255 (i.e., on an 8\ bit scale allowing for $2^8 = 256$ different values). 

3. by HEX\ values: 
Every color (or hue) is expressed by a character string (with the prefix\ `#`), followed by a triple of hexadecimal values (each ranging from\ `00` to\ `FF`) that correspond to the triple of decimal RGB\ values (in the range from\ 0 to\ 255).

<!-- Note: 3 ways to express the same color -->

Importantly, the _same_ color (or hue) can be denoted by its name, its RGB\ values, and its HEX\ values. 
Thus, the three different systems (i.e., color name, RGB values, or HEX values) can be _alternative_ representations of the same object. However, just as the properties of an alphabet or numeral notation system constraints the words or numbers that can be expressed by it, the universe of possible colors is limited by their representation. 
For instance, whereas the RGB and HEX systems usually distinguish between more than 16\ million (i.e., $(2^8)^3 = 256^3 = 16.777.216$) possible colors, only a limited number of them are named. Similarly, when multiple names refer to the same triple of RGB or HEX values, they denote the same color.^[Strictly speaking, different shades of a given color can be further distinguished by different _opacity_ or _transparency_ levels (typically denoted by an\ `alpha` value). However, the visual appearance of a transparent color can be mimicked by mixing the color with the background color (usually "white"). Hence, transparent shades of a color are usually not counted as different colors.]

<!-- Example: Inspecting and re-defining MPG_pal -->

<!-- Viewing colors (with seecol()):  -->

An example helps clarifying these abstract terms. 
When viewing an existing color or color palette with the `seecol()` function of **unikn**, the name, HEX and RGB values of the individual colors are being shown. 
For instance, the color palette `mpg_pal` provides the colors of the [Max Planck Society](https://www.mpg.de/en) as: 

```{r def-col-mpg-example, fig.width = 5, fig.asp = 1, fig.align = 'center'}
seecol(pal = unikn::mpg_pal, col_brd = "black", lwd_brd = .5)
```

As the middle color of `mpg_pal` is one of the `r length(colors())` named colors of `colors()`, R will recognize it as a color when we type its name (as a character object `"white"`) in a context that expects color inputs. 
However, the two other colors of `mpg_pal` have specifically been defined by the [Max Planck Society](https://www.mpg.de/en) and thus are _not_ pre-defined colors of R (i.e., not contained in `colors()`). 
To use such colors in\ R, we first need to define them as color objects. 

<!-- Defining colors in R: -->

So how can we define new colors in R? 
Most institutions will provide definitions of their corporate colors or instructions how they can be defined in some standard color system. 
Unfortunately, some definitions (such as those using the subtractive CMYK model) are aimed at printing (as opposed to computer screens) and other standards (such as numeric PANTONE&trade; PMS codes) are proprietary. 
But as most institutions have an interest in widely disseminating their colors, they typically include translations or approximations for alternative color models in their definitions. 

<a href="https://www.mpg.de/en">
<img src = "../inst/pix/logo_MPG.png" align = "right" alt = "MPG" width = "150px" style = "border:10;"/>
</a>

The [Max Planck Society](https://www.mpg.de/en) defines two main colors in its CD\ manual (available [here](https://docplayer.org/2328711-Max-planck-institut-das-erscheinungsbild-der-max-planck-gesellschaft-4-ueberarbeitete-auflage.html)): 

- A dark shade of green is defined as Pantone\ 328, and approximated by CMYK\ 100/0/57/30 or RGB\ 17/102/86. 

- A light shade of grey is Pantone\ 427, and approximated by CMYK\ 0/0/6/15 or RGB\ 221/222/214. 

In the following, we will first use the RGB values to define both colors (as these are provided in the definition).
However, we then show that we alternatively could have used the colors' HEX\ values (had they been provided) to define the same color objects.

#### Defining colors from their RGB values {-}

As our two colors of interest are given in RGB values, we can use the `rgb()` function of the **grDevices** package to define corresponding objects:

```{r def-col-rgb}
mpg_green <- grDevices::rgb( 17, 102,  86, maxColorValue = 255)
mpg_grey  <- grDevices::rgb(221, 222, 214, maxColorValue = 255)
```

The first three arguments specify values on the `red`, `green`, and `blue` dimension, respectively. 
Importantly, the values must be in the range\ $[0, max]$, where $max$ is provided by `maxColorValue = 255`. 
Whereas the [sRGB standard](https://en.wikipedia.org/wiki/SRGB) expresses all values in a normalized range of\ $[0, 1]$, most color definitions use the traditional 8-bit representation that distinguish between $2^8 = 256$\ levels in the range\ $[0, 255]$. 
The `rgb()` function additionally allows for adding\ `names` and a transparency level\ `alpha`, but as we can modify these aspects later, we do not need to use them at this point. 

What do we obtain when we evaluate our new color objects? 
Let's see: 

```{r def-col-rgb-eval}
mpg_green
mpg_grey
```

When evaluating an object created by `rgb()`, R returns a character sequence of the form\ `"#rrggbb"`. 
In this string, the sequence\ `rrggbb` expresses the three RGB values (for the `red`, `green`, and `blue` dimension, respectively) in hexadecimal (HEX) notation (in the range $[0, 255]$). 
In HEX\ notation, the ten numeral symbols (i.e., the digits from\ 0 to\ 9) or the decimal system are extended by the first six letters of the alphabet (i.e., A to\ F) to turn our familiar base-10 notation into an analogous place-value notation with a base of\ 16. 
Thus, the HEX\ values of `"#116656"` is merely a more compact way of expressing 

- `red`: $1 \cdot 16 + 1 = 17$
- `green`: $6 \cdot 16 + 6 = 102$
- `blue`: $5 \cdot 16 + 6 = 86$

Thus, the HEX values\ `"#116656"` provide an alternative way of expressing the RGB values\ 17/102/86 ---\ and both refer to the color "MPG green" (shown as `mpg_pal[1]` above). 


#### Defining colors from their HEX values {-}

The equivalence of the RGB and HEX notations can be illustrated by defining the same colors a second time, but this time specifying their HEX values (returned above):

```{r def-col-hex}
mpg_green_2 <- "#116656"
mpg_grey_2  <- "#DDDED6"
```

If both definitions really yield identical R objects, the following checks must evaluate to `TRUE`:

```{r def-col-rgb-hex-eual}
all.equal(mpg_green, mpg_green_2)
all.equal(mpg_grey,  mpg_grey_2)
```

Thus, we can define colors from both their RGB or HEX values ---\ and the two notations are really just two alternative ways of representing the same colors. 


### 2. Combining colors into a palette 

Having defined two new colors, we can easily combine them into a new color palette. 
A _color palette_ is simply a sequence of colors that are stored in some sequential data structure. 
The two simplest data structures in\ R that are suited for this purpose are vectors or data frames. 
Thus, the following two expressions both define valid color palettes: 

```{r cols-to-pals}
c(mpg_green, mpg_grey)
data.frame(mpg_green, mpg_grey)
```

As it often is easier to remember and access colors by their names, it usually makes sense to name the elements of a color palette (e.g., by using the `name()` function). 

<!-- Using newpal(): -->

The `newpal()` function of **unikn** provides a convenient wrapper for defining color palettes. 
Its main argument `col` uses a vector of color objects that make up the new palette, whereas `names` and `as_df` allow specifying color names and switching between returning the new palette as a vector or data base.    
Importantly, the color objects of `col` can be provided as RGB\ values, as R\ color names, or as HEX\ values:

```{r newpal-mpg-3, fig.width = 5, fig.asp = 1, fig.align = 'center'}
mpg_pal_3 <- newpal(col = c(rgb( 17, 102,  86, maxColorValue = 255), "white", "#DDDED6"),
                    names = c("MPG green", "white", "MPG grey"),
                    as_df = FALSE)
```

The new color palette `mpg_pal_3` contains the three basic colors of `mpg_pal` (shown above):

```{r seecol-mpg-3, fig.width = 5, fig.asp = 1, fig.align = 'center'}
seecol(mpg_pal_3, 
       main = "My new color palette", col_brd = "black", lwd_brd = .5)
```

The 5-color palette provided as `mpg_pal` includes two intermediate colors (at positions `mpg_pal[2]` and `mpg_pal[4]`) that are extrapolated by mixing the two main MPG colors with "white" in a balanced (50:50) fashion. 
The `usecol()` function of **unikn** automatically performs these extrapolations when extending a 3-element color palette (e.g., `mpg_pal_3`) to `n = 5` colors:

```{r newpal-mpg-5}
my_mpg_pal <- usecol(mpg_pal_3, n = 5)
```

The resulting color palette `my_mpg_pal` recreates the colors of the **unikn** color palette `mpg_pal` (shown above):

```{r seecol-mpg-5, fig.width = 5, fig.asp = 1, fig.align = 'center'}
seecol(my_mpg_pal, main = "My new MPG color palette", 
       col_brd = "black", lwd_brd = .5)
```

Technically, the two color palettes differ in some names and the way in which its middle color ("white") was specified (by its R color name or HEX\ values). 
However, when using `col2rgb()` for converting both palettes into their RGB\ values, we see that they contain the same colors:

```{r verify-equality}
all.equal(col2rgb(mpg_pal), col2rgb(my_mpg_pal), check.attributes = FALSE)
```

This concludes our illustration of defining a new colors and color palettes. 


## Comparing color palettes 

Once a new palette has been defined (and is available in your current R environment), the `seecol()` and `usecol()` functions allow comparing, scaling, and using them in the same fashion as the native **unikn** color palettes: 

```{r compare-custom-pals, eval = TRUE, fig.width = 5, fig.asp = .55, fig.align = 'center', collapse = TRUE}
# Re-define an existing color palette:
lmu_pal_rev <- newpal(col = lmu_pal_1[c(1, 3, 2)])

# Scaled version: 
seecol(list(lmu_pal_rev, uni_princeton_1, mpg_pal, pal_unikn), n = 7,
       pal_names = c("LMU Munich", "Princeton Uni", "Max Planck", "Uni Konstanz"),
       main = "Comparing scaled color palettes")
```

See the vignette on [Color recipes](color_recipes.html) for additional examples of using **unikn** functions for solving color-related tasks. 



## Resources

The following versions of **unikn** and corresponding resources are currently available: 

Type:                    | Version:           | URL:                           |         
:------------------------|:-------------------|:-------------------------------| 
A. **unikn** (R package): | [Release version](https://CRAN.R-project.org/package=unikn) | <https://CRAN.R-project.org/package=unikn> |
    &nbsp;                | [Development version](https://github.com/hneth/unikn/)      | <https://github.com/hneth/unikn/> | 
B. Online documentation:  | [Release version](https://hneth.github.io/unikn/)           | <https://hneth.github.io/unikn/> | 
    &nbsp;                | [Development version](https://hneth.github.io/unikn/dev/)   | <https://hneth.github.io/unikn/dev/> | 


## Vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a> 

The following vignettes provide instructions and examples for using the **unikn** colors, color palettes, and functions: 

<!-- Table of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color functions |
| 2. | [Color recipes](color_recipes.html) | Recipes for color-related tasks |
| 3. | [Institutional colors](color_inst.html) | Creating color palettes for other institutions |
| 4. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->