---
title: "Colors and color palettes"
author: "SPDS, uni.kn"
date: "2019 04 06"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(max.print = "75")
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               collapse = TRUE, 
               comment = "#>",
               message = FALSE,
               warning = FALSE,
               ## Default figure options:
               # fig.width = 7.2, 
               # fig.asp = .618, # golden ratio
               # out.width = "75%",
               fig.align = "center"
               )
```

Explain which colors and color palettes exist in `unikn` and how to use them.

Please load the `unikn` package to get started: 

```{r load_pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```

## Viewing color palettes

The function `seecol` is a general-purpose tool for viewing (or _seeing_) color palettes. 
`seecol` takes 2 main arguments:

1. `pal` describes either 1 or multiple color palettes (with a default `p = "all"`)
2. `n` describes the number of desired colors (with a default `n = "all"`).

Based on the setting of `pal`, the `seecol` function distinguishes between 2 modes: 


### Comparing multiple color palettes

When the `pal` argument specifies (a list of) multiple color palettes, `seecol` plots a vector for each palette to allow comparing these palettes. 
Some special keywords within the `unikn` package are `"all"`, `"unikn_basic"` and `grad_all"`. 

Viewing _all_ available color palettes:

```{r seecol_all, fig.width = 7, fig.height = 5}
seecol()  # same as seecol("all")
```

The set of `"all"` color palettes in `unikn` is sub-divided into 2 categories, which can be accessed by their names:

1. All _basic_ color palettes:

```{r seecol_unikn_all, fig.width = 7, fig.height = 3}
seecol("unikn_basic")
```

2. all pre-defined color _gradients_:

```{r seecol_grad_all, fig.width = 7, fig.height = 4}
seecol("grad_all")
```


### Viewing a particular color palette 

When the `pal` argument specifies a single color palette, `seecol` plots a more detailed view of this particular color palette:

```{r view_pal, fig.width = 7, fig.asp = .618}
seecol(pal_unikn)
```

Note that the `seecol` function is quite permissive with respect to specifying its `pal` argument: 
A particular color palette (e.g., `pal_seegruen`) can not only be displayed by providing it (as an object) but also by providing its name (i.e., `"pal_seegruen"`) or even its incomplete object name or name (i.e., `"seegruen"` or `seegruen`). 
Hence, the following all yield the same result: 

```{r, eval = FALSE}
seecol(pal_seegruen)
seecol("pal_seegruen")
seecol(seegruen)  # issues a warning
seecol("seegruen")
```

### Extending and mixing color palettes

__ToDo__ (see examples below)


## Using colors and color palettes 

The `usecol` function corresponds to `seecol` by taking the same 2 main arguments (`pal` and `n`). 
However, as its purpose is _using_ the colors specified by `pal`, rather than plotting (or _seeing_) them, the `pal` argument typically contains only one color palette. 

### Using an existing color palette

Using an exising color palette in a graph:

```{r use_col_default, fig.width = 7, fig.asp = .618}
# Using a color palette:
barplot(1/sqrt(1:11), col = usecol(pal_unikn))
```



Alternatively we can use `seecol` to view, save and subesequently use a color palette in a graph: 

We can first view a color palette (to _see_ whether we like it):  

```{r save_see_my_pal, fig.width = 7, fig.height = 6}
my_pal <- seecol(pal_unikn_pair)  # save and view the palette. 
```

and then _use_ it (here: the colors saved as `my_pal`) in a plot: 

```{r use_my_pal, fig.width = 7, fig.height = 4}
barplot(matrix(sample(5:10, size = 16, replace = TRUE), ncol = 8), beside = TRUE, col = my_pal)
```


## Customizing color palettes

Both `seecol` and `usecol` allow a flexible on-the-fly customization of color palettes.

### Selecting subsets

Using only a subset of colors:

```{r use_subset, fig.width = 7, fig.height = 6}
seecol("all", n = 4)
seecol(pal_unikn, 4)
```

Importantly, when using pre-defined color palettes of `unikn` but a value of `n` that is smaller than the length of the current color palette, `usecol` and `seecol` select a reasonable subset of colors:

```{r use_col_small_n, fig.width = 4, fig.asp = .618}
barplot(1/sqrt(1:2), col = usecol(pal_seeblau, n = 2))
barplot(1/sqrt(1:3), col = usecol(pal_seeblau, n = 3))
```

### Extending color palettes

For values of `n` that are larger than the number of available colors in `pal`, the specified color palette is extended:

```{r use_superset, fig.width = 7, fig.height = 6}
seecol("all", n = 12)
seecol(pal_seeblau, 12)
```

When using a color palette: 

```{r use_col_big_n, fig.width = 7, fig.asp = .618}
barplot(1/sqrt(1:11), col = usecol(pal_bordeaux, n = 11))
```

### Mixing color palettes

Combining colors and color palettes: 

```{r combine_pals, fig.width = 7, fig.asp = .618}
seecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
seecol(pal = c(rev(pal_petrol), "white", pal_bordeaux))
```

Combining colors and color palettes to get a specific number of colors: 

```{r combine_pals_n, fig.width = 7, fig.asp = .618}
seecol(pal = c(Karpfenblau, Seeblau, "gold"), n = 10)          # no names
seecol(pal = usecol(c(Karpfenblau, Seeblau, "gold"), n = 10))  # shows HEX values as names
```

Examples of _using_ combinations of colors and color palettes to create new color palettes:

```{r bar_seeblau_pinky, fig.width = 7, fig.asp = .618}
barplot(1/sqrt(1:11), col = usecol(c(rev(pal_seegruen), "white", pal_bordeaux), n = 11))
barplot((-5:5)^2+5,   col = usecol(pal = c(rev(pal_seeblau), "white", pal_pinky)))
barplot(1/sqrt(11:1), col = usecol(pal = c(rev(pal_seeblau), "white", pal_seegruen)))
```


### Comparing custom palettes 

Advanced users can also use `seecol` for displaying and comparing different custom palette. The `pal` argument is a list of colors or color palettes:

```{r compare_custom_palettes, fig.width = 7, fig.height = 3}
pal1 <- c(rev(pal_seeblau), "white", pal_pinky)
pal2 <- usecol(c(Karpfenblau, Seeblau, "gold"), n = 10)

seecol(list(pal1, pal2, pal_unikn, pal_unikn_pair))
# TODO: Does currently not work for unikn_pair, unikn_ppt, as no defaults are defined there. 
```


## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color palettes | 
| 2. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->
