---
title: "Colors and color palettes"
author: "SPDS, uni.kn"
date: "2019 03 23"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(max.print = "75")
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               collapse = TRUE, 
               comment = "#>",
               message = FALSE,
               warning = FALSE,
               ## Default figure options:
               # fig.width = 7.2, 
               # fig.asp = .618, # golden ratio
               # out.width = "75%",
               fig.align = "center"
               )
```

Explain which colors and color palettes exist in `unikn` and how to use them.

Please load the `unikn` package to get started: 

```{r load_pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```

## Viewing color palettes

Distinguish between 2 main modes: 

1. Viewing all available color palettes:

```{r seecol_all, fig.width = 7, fig.height = 6}
seecol()  # same as seecol("all")
```

There are 2 subsets of the available color palettes: 

a. All basic color palettes:

```{r seecol_unikn_all, fig.width = 7, fig.height = 3}
seecol("unikn_all")
```

b. all pre-defined color gradients:

```{r seecol_grad_all, fig.width = 7, fig.height = 3}
seecol("grad_all")
```

2. Viewing the colors and details of a particular color palette:

```{r view_pal, fig.width = 7, fig.asp = .618}
seecol(pal_unikn_plus)
```

Note that the `seecol` function is quite permissive with respect to specifying its `pal` argument: 
A particular color palette (e.g., `pal_seegruen`) can not only be displayed by providing it (as an object) but also by providing its name (i.e., `"pal_seegruen"`) or even its incomplete object name or name (i.e., `"seegruen"` or `seegruen`). 
Hence, the following all yield the same result: 

```{r, eval = FALSE}
# TODO: Run them all? 
seecol(pal_seegruen)
seecol(seegruen)
seecol("pal_seegruen")
seecol("seegruen")
```

## Using colors and color palettes with `usecol`

Using colors in a graph:

```{r use_col, fig.width = 7, fig.asp = .618}
# Using a color palette:
barplot(1/sqrt(1:11), col = usecol(pal_unikn_plus))

barplot(1/sqrt(1:11), col = usecol(pal_bordeaux, n = 11))
barplot(1/sqrt(1:11), col = usecol(c(rev(pal_seegruen), "white", pal_bordeaux), n = 11))
```

## Modifying color palettes

Using only a subset of colors:

```{r use_subset, fig.width = 7, fig.height = 6}
seecol("all", n = 4)
seecol(pal_unikn_plus, 4)
```

Extending a palette to a larger number of colors than contained in the palette:

```{r use_superset, fig.width = 7, fig.height = 6}
seecol("all", n = 12)
seecol(pal_seeblau, 12)
```

## Mixing color palettes

Combining colors and color palettes: 

```{r combine_pals, fig.width = 7, fig.asp = .618}
seecol(pal = c(rev(pal_seeblau), "white", pal_pinky))
seecol(pal = c(rev(pal_petrol), "white", pal_bordeaux))
```

Combining colors and color palettes to get a specific number of colors: 

```{r combine_pals_n, fig.width = 7, fig.asp = .618}
seecol(pal = c(Karpfenblau, Seeblau, "gold"), n = 10)          # no names
seecol(pal = usecol(c(Karpfenblau, Seeblau, "gold"), n = 10))  # shows HEX values as names
```

Examples of combining colors and color palettes into new color palettes:

```{r bar_seeblau_pinky, fig.width = 7, fig.asp = .618}
barplot((-5:5)^2+5, col = mixcol(pal = c(rev(pal_seeblau), "white", pal_pinky)))
barplot(1/sqrt(11:1), col = mixcol(pal = c(rev(pal_seeblau), "white", pal_seegruen)))
```


## Comparing custom palettes 

With `seecol` you can also disply different custom palettes for comparison, when passign the argument as a list:

```{r}
pal1 <- c(rev(pal_seeblau), "white", pal_pinky)
pal2 <- usecol(c(Karpfenblau, Seeblau, "gold"), n = 10)

seecol(list(pal1, pal2, pal_unikn, pal_unikn_pair))
# TODO: Does currently not work for unikn_pair, unikn_ppt, as no defaults are defined there. 
```


## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color palettes | 
| 2. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->
