---
title: "Institutional colors"
author: "SPDS, uni.kn"
date: "2022 12 05"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Institutional colors}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
options(max.print = "75")
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               collapse = TRUE, 
               comment = "#>",
               message = FALSE,
               warning = FALSE,
               ## Default figure options:
               # fig.width = 6, 
               # fig.asp = .8 # .618, # golden ratio
               fig.align = "center",
	             out.width = "60%"
               )
```

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

The color palettes included in the **unikn** package are based on the CD\ manual of the [University of Konstanz](https://www.uni-konstanz.de/). 
However, the functionality provided by the package makes it easy and straightforward to define and use **your own colors**.  

This vignette first shows which color palettes from other institutions are currently provided by the **unikn** package. 
As these will never be complete, it then explains how the color-related tools of **unikn** can be used for **creating new color palettes**. 
<!-- Main color info: --> 
(For a general introduction to the package's colors, color palettes, and color functions, see the vignettes on [Colors and color functions](colors.html) and [Color recipes](color_recipes.Rmd). 
The vignette on [text](text.html) provides information on creating text boxes and decorations.) 

We install or load the **unikn** package to get started: 

```{r load-pkg-colors, message = FALSE, warning = FALSE}
# install.packages('unikn')  # install unikn from CRAN client
library('unikn')             # load the package
```


## Current color palettes

### Overview

In addition to the specific color palettes of the [University of Konstanz](https://www.uni-konstanz.de/), 
**unikn** currently provides the general color palettes of the following institutions: 

- [Caltech](https://www.caltech.edu/): `caltech_pal_1`, `caltech_pal_2`, `caltech_pal_3` 
- [ETH Zurich](https://ethz.ch/): `eth_pal`, `eth_pal_light`, `eth_pal_grey` 
- [Free University Berlin](https://www.fu-berlin.de/en/index.html): `fu_pal_0`, `fu_pal_1`, `fu_pal_2`, `fu_pal_3` 
- [Humboldt University Berlin](https://www.hu-berlin.de/en): `hu_pal_1`, `hu_pal_2` 
- [Max Planck Society](https://www.mpg.de/en): `mpg_pal` 
- [University of Freiburg](https://uni-freiburg.de/): `uni_freiburg_br`, `uni_freiburg_blue`, `uni_freiburg_grey`, `uni_freiburg_info` 
- [University of Göttingen](https://uni-goettingen.de/): `uni_goettingen_1`, `uni_goettingen_2`, `uni_goettingen_3` 
- [University of Hamburg](https://www.uni-hamburg.de/): `uni_hamburg_1`, `uni_hamburg_2` 
- [University of Köln](https://www.uni-koeln.de/): `uni_koeln_1`, `uni_koeln_2` 
- [University of Konstanz](https://www.uni-konstanz.de/): `uni_konstanz`, `uni_konstanz_pref` 
- [Princeton University](https://www.princeton.edu/): `uni_princeton_0`, `uni_princeton_1`, `uni_princeton_2` 
- [University of Regensburg](https://www.uni-regensburg.de/): `uni_regensburg_1`, `uni_regensburg_2`, `uni_regensburg_3` 
- [Rensselaer Polytechnic Institute](https://www.rpi.edu/): `rpi_pal_1`, `rpi_pal_2`, `rpi_pal_3`

<!-- Note: Number of palettes & exported as named vectors: -->

These 38\ palettes are exported (as named vectors) and documented (to credit their contributors and sources). 
We can use and modify these palettes with the **unikn** [color functions](colors.html):

<!-- Note: Duplicate image in README.Rmd -->

```{r show-add-pals, fig.width = 7.2, out.width = "550px", fig.asp = 1.5}
seecol("add", main = "Color palettes from various institutions")
```


### Call for contributions

<!-- Collecting contributed color palettes: -->

> **Collecting color palettes**
>
> Are you using **unikn** to create **your own color palettes**?  
> If so, please **let us know** (e.g., by posting on this [GitHub issue](https://github.com/hneth/unikn/issues/11))!  
>  
> We collect color palettes from other institutions to make them available in future versions of **unikn**.  
>   
> To enable us to verify and provide appropriate credit to your contributions, please provide us with the following information:  
>   
> 1. your code (e.g., the `newpal()` command creating your color palette),   
> 2. your source (e.g., some URL with color definitions),   
> 3. your name and some contact information (e.g., an Email address).  
>  
> We're looking forward to your inputs and contributions!  


## Defining colors and color palettes 

The following examples show how to define color palettes (with dedicated color names) for three institutions. 
They can be easily adapted to any other group, company, or entity, that uses some pre-defined set of colors. 

```{r Cardiff, echo = FALSE, eval = FALSE, fig.align = 'center', fig.width = 5, fig.height = 3}
# Cardiff University, UK: 
# ToDo. 
```

### A. [University of Freiburg](https://uni-freiburg.de/), Germany: 
<a href="https://uni-freiburg.de/">
<img src = "../inst/pix/logo_freiburg.png" align = "right" width = "120px" alt = "Uni Freiburg" style = "border:10;"/> 
</a>

The [Albert-Ludwigs Universität Freiburg](https://uni-freiburg.de/) provides fairly extensive information on its corporate color scheme (available at `portal.uni-freiburg.de/cmsforum/wsg/webstyleguide/farben` on 2022-10-20). 
Colors can be defined in a variety of ways, but R comes with convenient tools (like the `col2rgb()` and `convertColor()` functions of **grDevices**) to handle most cases. 
The most straightforward way of creating a new color palette in R is by using its HEX/HTML code (provided in character format). 
As the [University of Freiburg](https://uni-freiburg.de/) kindly provides their colors in this format, we can easily define the corresponding color palettes as named vectors: 

```{r freiburg-cols, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
# Source: http://portal.uni-freiburg.de/cmsforum/wsg/webstyleguide/farben

# Basic colors: ---- 

uni_freiburg_br <- c("#004a99", "#c1002a")
names(uni_freiburg_br) <- c("uni-blau", "uni-rot")

uni_freiburg_bwr <- c("#004a99", "white", "#c1002a")  # add "white" for better gradients
names(uni_freiburg_bwr) <- c("uni-blau", "weiss", "uni-rot")

# Web colors: ----

uni_freiburg_blue <- c("#004a99", "#2a6ebb", "#6f9ad3")
names(uni_freiburg_blue) <- c("blue-1", "blue-2", "blue-3")

uni_freiburg_grey <- c("#f2f3f1", "#e0e1dd", "#d5d6d2", "#c9cac8", 
                       "#b2b4b3", "#9a9b9c", "#747678", "#363534")
names(uni_freiburg_grey) <- c("grey-0", "grey-1", "grey-2", "grey-3", 
                              "grey-5", "grey-7", "grey-9", "grey-font")

# Visualizations / illustrations: ----

uni_freiburg_info <- c("#2a6ebb", "#a7c1e3", "#7b2927", "#de3831", "#739600", "#92d400", 
                       "#4d4f53", "#747678", "#b2b4b3", "#d5d6d2", "#e98300", "#efbd47")
names(uni_freiburg_info) <- c("mid-blau", "hell-blau", "dark-red", "hell-red", "mid-green", "hell-green", 
                              "anthrazit", "dark-grey", "mid-grey", "hell-grey", "orange", "gelb")
```

Alternatively, we can define both colors and names in one step by using the `newpal()` function:

```{r freiburg-newpal}
uni_freiburg_info <- newpal(col = c("#2a6ebb", "#a7c1e3", "#7b2927", "#de3831", "#739600", "#92d400", 
                                    "#4d4f53", "#747678", "#b2b4b3", "#d5d6d2", "#e98300", "#efbd47"),
                            names = c("mid-blau", "hell-blau", "dark-red", "hell-red", "mid-green", "hell-green", 
                                      "anthrazit", "dark-grey", "mid-grey", "hell-grey", "orange", "gelb")
                            )
```

Using `newpal()` makes it easy to combine colors and their names and return the resulting color palette in either data frame or vector format.

Once a new color palette has been defined (in your current R\ environment), use all color-related functions to view, modify, and use the palette: 

```{r freiburg-seecol, fig.width = 5, fig.asp = .80, fig.align = 'center', collapse = TRUE}
# View color palette:
seecol(uni_freiburg_info, main = "Colors of the University of Freiburg")
# seecol(uni_freiburg_bwr, n = 7)  # extend color palette
# seecol(c(uni_freiburg_blue, "white", uni_freiburg_grey))  # mix color palette

# use in a demo plot:
demopal(uni_freiburg_info,
        type = 4, col_par = NA, seed = 79098, 
        main = "Colors of the University of Freiburg")
```


### B. [Princeton University](https://princeton.edu/), USA: 
<a href="https://princeton.edu/">
<img src = "../inst/pix/logo_princeton.jpg" align = "right" width = "180px" alt = "Princeton University" style = "border:10;"/>
</a>

The color scheme of [Princeton University](https://princeton.edu/) is easily recognized by its combination of orange with black and white elements. 
The official guidelines (available [here](https://communications.princeton.edu/guides-tools/logo-graphic-identity)) define "Princeton Orange" as Pantone (PMS) 158\ C. 

- The PANTONE&trade; color finder at <https://www.pantone.com/connect/158-C> yields the following color values:

    - RGB: `232 119 34`  
    - HEX/HTML: `#E87722`  
    - CMYK: `0 62 95 0`  

- However, the guide also specifies and distinguishes between two additional versions of orange (depending on background color) and provides the following HEX/HTML values for them:  

    - Orange on white: "#E77500"  
    - Orange on black: "#F58025"  

These definitions suggest defining three separate versions of orange and corresponding color palettes: 

```{r princeton-1, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
# 3 shades of orange: 
orange_basic <- "#E87722"  # Pantone 158 C
orange_white <- "#E77500"  # orange on white background
orange_black <- "#F58025"  # orange on black background

# Defining color palettes:
uni_princeton <- c(orange_basic, "black")
names(uni_princeton) <- c("orange", "black")

# Define color palette and names (in 2 steps):
uni_princeton_1 <- c(orange_white, "white", "black")
names(uni_princeton_1) <- c("orange_w", "white", "black")
```

Again, we can define a palette of named colors in one step:

```{r princeton-2, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
# Basic colors:
uni_princeton_0 <- newpal(col = c(orange_basic, "black"),
                          names = c("orange", "black"))

# 1st color palette (on white backgrounds):
uni_princeton_1 <- newpal(col = c(orange_white, "white", "black"),
                          names = c("orange_w", "white", "black"))
                          
# 2nd color palette (on black backgrounds):
uni_princeton_2 <- newpal(col = c(orange_black, "black", "white"), 
                          names = c("orange_b", "black", "white"))
```

and then inspect a new color palette (with a custom background color and title):

```{r princeton-3, fig.width = 5, fig.asp = .80, fig.align = 'center', fig.show = "hold", collapse = TRUE}
seecol(uni_princeton_1, col_bg = "grey90", 
       main = "Colors of Princeton University")
```


### C. [Max Planck Society](https://www.mpg.de/en), Germany: 
<a href="https://www.mpg.de/en">
<img src = "../inst/pix/logo_MPG.png" align = "right" alt = "MPG" width = "150px" style = "border:10;"/>
</a>

The CD manual (available [here](https://docplayer.org/2328711-Max-planck-institut-das-erscheinungsbild-der-max-planck-gesellschaft-4-ueberarbeitete-auflage.html)) of the [Max Planck Society](https://www.mpg.de/en) specifies the use of two primary colors: 

- Green as Pantone 328: Using the PANTONE&trade; color finder at <https://www.pantone.com/connect/328-C> yields the following color values:

    - RGB: `0 115 103`    
    - HEX/HTML: `#007367`   
    - CMYK: `100 10 61 38`  

- Grey as Pantone 427: Using the PANTONE&trade; color finder at <https://www.pantone.com/connect/427-C> yields the following color values: 

    - RGB: `208 211 212`  
    - HEX/HTML: `#D0D3D4`   
    - CMYK: `7 3 5 8`   

```{r mpg-rgb-cols, echo = FALSE, eval = FALSE, fig.width = 6, fig.asp = .618, fig.align = 'center', collapse = TRUE}
# Using RGB values: 
mpg_pal <- data.frame(# named colors: 
  "mpg_green" = rgb( 17, 102,  86, maxColorValue = 255),
  "mpg_gray"  = rgb(221, 222, 214, maxColorValue = 255),
  "mpg_grey"  = rgb(208, 211, 212, maxColorValue = 255)
  )
seecol(pal = list(mpg_pal))  # Ooops...
```

```{r mpg-hex-cols, echo = FALSE, eval = FALSE}
# Using HEX values:
mpg_green <- "#007367"
mpg_grey  <- "#D0D3D4"

# col2rgb(mpg_green)
# col2rgb(mpg_grey)

mpg_pal <- usecol(pal = c(mpg_green, "white", mpg_grey))
# seecol(mpg_pal)
```

Again, the easiest way of defining a corresponding color palette is by defining the elementary colors and then 
combine them into a named color palette by using the `newpal()` function. 
To enable color gradients, we insert the color\ `"white"` between the two\ dedicated colors: 

```{r mpg-newpal}
# Main colors (using HEX values):
mpg_green <- "#007367"
mpg_grey  <- "#D0D3D4"

# Create new color palette (as a vector):
mpg_pal <- newpal(col = c(mpg_green, "white", mpg_grey), 
                  names = c("mpg green", "white", "mpg grey")
                  )
```

As before, can now use the `seecol()` and `usecol()` functions to view, modify, and use the new `mpg_pal` color palette: 

```{r mpg-seecol, fig.width = 5, fig.asp = .80, fig.align = 'center', collapse = TRUE}
seecol(mpg_pal,
       col_brd = "black", lwd_brd = 1, 
       main = "Colors of the Max Planck Society")
```


## Comparing custom color palettes 

Once a new palette has been defined (and is available in your current R environment), the `seecol()` and `usecol()` functions allow comparing, scaling, and using them in the same fashion as the native **unikn** color palettes: 

```{r compare-custom-pals, eval = TRUE, fig.width = 5, fig.asp = .50, fig.align = 'center', collapse = TRUE}
# Compare and extend custom color palettes:
# seecol(list(uni_freiburg_info, uni_freiburg_bwr, mpg_pal, uni_princeton_1))

# Scaled version: 
seecol(list(uni_freiburg_bwr, uni_princeton_1, mpg_pal, pal_unikn), n = 9,
       pal_names = c("Uni Freiburg", "Princeton Uni", "Max Planck", "Uni Konstanz"),
       main = "Comparing scaled color palettes")
```


## Resources

The following versions of **unikn** and corresponding resources are currently available: 

Type:                    | Version:           | URL:                           |         
:------------------------|:-------------------|:-------------------------------| 
A. **unikn** (R package): | [Release version](https://CRAN.R-project.org/package=unikn) | <https://CRAN.R-project.org/package=unikn> |
    &nbsp;                | [Development version](https://github.com/hneth/unikn/)      | <https://github.com/hneth/unikn/> | 
B. Online documentation:  | [Release version](https://hneth.github.io/unikn/)           | <https://hneth.github.io/unikn/> | 
    &nbsp;                | [Development version](https://hneth.github.io/unikn/dev/)   | <https://hneth.github.io/unikn/dev/> | 


## Vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a> 

The following vignettes provide instructions and examples for using the **unikn** colors, color palettes, and functions: 

<!-- Table of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color functions |
| 2. | [Color recipes](color_recipes.html) | Recipes for color-related tasks |
| 3. | [Institutional colors](color_inst.html) | Creating color palettes for other institutions |
| 4. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->