---
title: "Colors and color palettes"
author: "Hansjörg Neth, SPDS, uni.kn"
date: "2019 03 05"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Explain which colors and color palettes exist in `unikn` and how to use them.

Please load the `unikn` package to get started: 

```{r load_pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```


## Viewing color palettes

Viewing all available color palettes:

```{r seepal_all, fig.align = 'center', fig.width = 7.2, fig.height = 6}
seepal()
```

Viewing the colors and details of a particular color palette:

```{r view_pal, fig.align = 'center', fig.width = 7.2, fig.height = 5}
seepal(pal = pal_unikn_plus)
```

## Using colors and color palettes

Using colors in a graph:

```{r}
# Using a color palette:
barplot(1/sqrt(1:11), col = seepal(pal_unikn_plus))
```





## Modifying color palettes

Using only a subset of colors:

```{r, fig.align = 'center', fig.width = 6, fig.height = 5}
seepal(n = 4)
seepal(pal_unikn_plus, 4)
```

Extending a palette to a larger number of colors than contained in the palette:

```{r}
# seepal(n = 12)
# seepal(pal_unikn_plus, 12)
```

Combining color palettes: 

```{r, fig.align = 'center', fig.width = 7.2, fig.height = 5}
seepal(pal = c(rev(pal_seeblau), "white", pal_petrol))
seepal(pal = c(rev(pal_petrol), "white", pal_bordeaux))
```

Mixing palettes:

```{r}
seepal(pal = col_scale(c(Karpfenblau, Seeblau, "gold"))(10))
```

Examples of combining colors and color palettes into new color palettes:

```{r bar_seeblau_pinky, fig.align = 'center', fig.width = 7.2, fig.height = 4}
barplot((-5:5)^2+5, col = seepal(pal = c(rev(pal_seeblau), "white", pal_pinky)))
barplot(1/sqrt(11:1), col = seepal(pal = c(rev(pal_seeblau), "white", pal_seegruen)))
```

## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color palettes | 
| 2. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->
