---
title: "Colors"
author: "Hansjörg Neth, SPDS, uni.kn"
date: "2019 03 03"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Colors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Explain which colors and color palettes exist in `unikn` and how to use them.

```{r}
library(unikn)
```


## Viewing color palettes

Viewing all available color palettes:

```{r}
seepal()
```

Viewing a particular color palette:

```{r}
seepal(pal = pal_unikn_plus)
```

## Using colors and color palettes

```{r}
seepal(pal_unikn_pref)
```

## Modifying color palettes

Using only a subset of colors:

```{r}
seepal(n = 4)
seepal(pal_unikn_plus, 4)
```

Extending a palette to a larger number of colors than contained in the palette:

```{r}
# seepal(n = 12)
# seepal(pal_unikn_plus, 12)
```

Combining color palettes: 

```{r}
seepal(pal = c(rev(pal_seeblau), "white", pal_pinky))
seepal(pal = c(rev(pal_seeblau), "white", pal_petrol))
seepal(pal = c(rev(pal_petrol), "white", pal_bordeaux))
```


## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/box_unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Using color palettes | 
| 2. | [Text](text.html)     | Using decorated text | 


<!-- eof. -->


<!-- Default vignette stuff: --> 

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
