---
title: "Text boxes and decorations"
author: "SPDS, uni.kn"
date: "2023 01 01"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Text boxes and decorations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(max.print = "75")

knitr::opts_chunk$set(echo = TRUE,
                      cache = FALSE,
                      collapse = TRUE, 
                      comment = "#>",
                      prompt = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      # Default figure options:
                      fig.align = "center" 
                      # fig.width = 6, 
                      # fig.asp = .8 # .618, # golden ratio
                      # out.width = "60%"
                      )

# URLs: ------ 

# unikn / Uni Konstanz:
url_unikn <- "https://www.uni-konstanz.de"

# unikn package:
url_unikn_cran   <- "https://CRAN.R-project.org/package=unikn"
url_unikn_github <- "https://github.com/hneth/unikn"

# unicol package: 
url_unicol_cran   <- "https://CRAN.R-project.org/package=unicol"
url_unicol_github <- "https://github.com/hneth/unicol"
```

<!-- unikn pkg logo and link: -->
<a href = "https://CRAN.R-project.org/package=unikn">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "155px" style = "width: 155px; float: right; border:0;"/>
</a>

This vignette explains the **unikn** functions for plotting graphical elements and text decorations. 
(See the vignette on [colors](colors.html) for general information on colors, color palettes, and corresponding functions, 
or the more specialized vignettes on [color recipes](color_recipes.html) and [institutional colors](color_inst.html).) 

Please install and/or load the **unikn** package to get started: 

```{r load-pkg-text, message = FALSE, warning = FALSE}
# install.packages('unikn')  # install unikn from CRAN client
library('unikn')             # loads the package
```

## Overview 

Beyond providing support for [colors](colors.html), **unikn** provides functions for plotting graphical elements (like boxes) and text decorations (with colored backgrounds or underlining). 
Before illustrating these functions, please note some caveats: 

- Plotting text (i.e., graphically rendering characters) is rarely a good idea. In contrast to the vector fonts used by DTP programs, plotted text typically does not scale (when changing the size of images) and cannot be recognized automatically (e.g., copied, indexed, or scraped). 
Hence, the following functions should only be used when no better solutions are available (e.g., for adding text to graphs, or for obstructing automatic text recognition). 

- Like all other templates, our renderings are subject to constraints and limitations. For instance, as any standard installation of [R](https://www.r-project.org/) lacks the prescribed "Theinhardt" fonts, we only mimic the official design specifications (in Arial, sans serif). 

That said, digital decorations can be useful to emphasize or structure information, provided that they can be produced easily and flexibly. The **unikn** package helps preventing common mistakes by novices (e.g., boxes or lines extending beyond text, or step-functions in multi-line headings) and can be customized and improved by expert users. Hence, we hope that the following functions are useful for plotting graphical elements (e.g., boxes, logos, etc.) and help you achieve a uniform look when styling presentations or visualizations. 

Please load the **unikn** package to get started: 

```{r load-pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```

## Plotting empty slides and boxes

As the text functions of **unikn** add styled text to an existing plot, we must first create a plot to which we can add text. 
**unikn** provides 2 basic functions for creating new plots:

1. `slide` plots an empty canvass (a rectangular box or frame); 
2. `xbox` plots a colored box with an X in its top right corner.

Here are both commands with their default settings: 

```{r slide-def, fig.align = 'center', fig.width = 3.5, fig.height = 2.5}
slide()  # plot default slide
```

```{r xbox-def, fig.align = 'center', fig.width = 2, fig.height = 2}
xbox()  # plot default xbox
```

Both functions accept additional arguments to customize their appearance. 

### Colors

The background and border colors of slides can be changed by setting their `col`, `border`, and `lwd` arguments:

```{r slide-col, fig.align = 'center', fig.width = 3.5, fig.height = 2.5}
slide(col = pal_seeblau[[1]], border = Petrol, lwd = 5)  # set colors of slide background and border
slide(col = pal_peach[[2]], border = NA)  # colored slide without a border
```

Similarly, the background color of xboxes can be changed by the `col` argument:

```{r xbox-col, fig.align = 'center', fig.width = 2, fig.height = 2}
xbox(col = Pinky)  # set background color of xbox
```

### Dimensions

By default, both `slide` and `xbox` fill the entire plotting region (i.e., all margins are set to 0).

The width of a slide is set to 4/3\ of its height (of\ 1). 
To change this default, use the `dim`\ argument for specifying the slide's x- and y-dimensions, respectively:

```{r slide-dim, fig.align = 'center', fig.width = 4.5, fig.height = 1.5}
slide(dim = c(3, 1))  # a slide 3x wider than high
```

Note that changes in slide dimensions are invisible when viewing images interactively, as R automatically scales images to the size of the canvass. They mostly affect the relation between metrics when adding elements to an\ object.

Similarly, an xbox has the default dimensions of a unit square (i.e., `dim = c(1, 1)`). 
To change this default, use the `dim` argument for specifying the xbox's `x`- and `y`-dimensions, respectively:

```{r xbox-dim, fig.align = 'center', fig.width = 1.5, fig.height = 3}
xbox(col = Petrol, dim  = c(1, 2))  # an xbox 2 times higher than wide
```

The two strokes of the\ X (in the top right corner) are orthogonal when viewing the `xbox` at the aspect ratio indicated by `dim` (i.e., here: at a ratio of width x = 1 to height y = 2). When viewing the box at any different aspect ratio, the X appears to be distorted. Hence, viewing and printing plots at the desired aspect ratio helps when creating an xbox.

## Plotting styled text

There are three main functions for plotting text with decorations: 

- `mark()` allows highlighting text with colored background boxes; 
- `uline()` allows underlining text with colored lines; 
- `post()` allows posting text in rectangular `xbox` elements.

In addition, there are 2 auxiliary functions: 

- `headline()` allows plotting headlines with colored background boxes; 
- `url_unikn()` allows formatting URLs the uni.kn way. 

The following sections illustrate the functions and their main options in turn. 

### Mark text

The `mark()` function allows highlighting text by plotting it with colored background boxes (to provide the functionality of "Markieren"): 

```{r mark-demo-label, echo = FALSE, fig.width = 3, fig.height = 3}
lbl_blank <- c("                                                ",
              "                                      ",
              "                                                      ",
              "                                                ",
              "                              ")

lbl_blank <- c(paste(rep(" ", 48), collapse = ""), 
               paste(rep(" ", 38), collapse = ""), 
               paste(rep(" ", 54), collapse = ""), 
               paste(rep(" ", 48), collapse = ""), 
               paste(rep(" ", 30), collapse = ""))
```

```{r mark-demo-fig, echo = FALSE, fig.width = 3, fig.height = 3}
# (a) Create a new plot (of type "slide"):
mark(labels = lbl_blank,
     x = 0, y = .85, y_layout = .03,
     col_bg = Seeblau,
     cex = 1.0,
     new_plot = "slide")
```

By default, `mark()` assumes that we want to add highlighted text to an\ existing plot. 
The text labels to be added are provided (as a character vector) by the `labels` argument, their coordinates as the corresponding\ `x` and\ `y` arguments. 
Various text decorations (i.e., text and background colors, text sizes, and font types) can be set by specifying additional arguments (i.e., `col`, `col_bg`, `cex`, and `font`, respectively): 

```{r set-par, echo = FALSE}
# Set par:
opar <- par(no.readonly = TRUE)  # all par settings that can be changed.
par(mar = c(1, 1, 1, 1))
```


```{r mark-messy-plot, echo = TRUE, fig.width = 5.5, fig.height = 3.5, out.width = "60%"}
# Some pre-existing plot: 
n <- 20
set.seed(1)
par(mar = c(0, 0, 0, 0))
plot(x = runif(n), y = runif(n), type = "p", pch = 16, cex = 20, col = grey(0, .2),
     axes = FALSE, xlab = "", ylab = "")

# Add 1 marked label:
mark(labels = "Some messy plot", x = .02, y = .9)

# Add 2 more marked labels:
mark(labels = c("Note something here.", "More highlighting here."), 
     x = c(.1, .6), y = c(.2, .4),
     col_bg = c(pal_seeblau[[2]], pal_peach[[3]]), cex = 1.2)
```

```{r reset-opar, echo = FALSE}
# Restore plot settings:
par(opar)
```

If no pre-existing plot is available, setting the `new_plot` argument to either\ `"blank"` or\ `"slide"` will generate a new default plot (by calling the `slide()` function described above):

```{r mark-new-plot, echo = TRUE, fig.width = 3.5, fig.height = 3.5, out.width = "35%"}
# A sentence to plot: 
s <- c("Markieren",
       "ist ein Bestandteil",
       "von Studieren.")

# Create a new plot (of type "slide"):
mark(labels = s,
     x = 0, y = .85, y_layout = .03,
     col_bg = Seeblau,
     cex = 1.5,
     new_plot = "slide")

# Add text to the same plot:
mark(labels = s,
     x = 0, y = .40, y_layout = .03,
     col_bg = pal_pinky[[2]],
     cex = 1.5, 
     new_plot = "none")
```

Rather than providing explicit `x`\ and `y`\ coordinates for all text elements in\ `labels`, the calls above defined an initial coordinate and used a numeric `y_layout` argument to fix the distance between consecutive text labels (to a value of\ `.03`). 

As an alternative to numeric coordinates and position values, `x_layout` and `y_layout` also accept some special values (as character strings):

1. `x_layout` accepts: 

    - `x_layout = "center"` for centering labels horizontally (on the first label or the middle),  
    - `x_layout = "left"` left-justifies labels (on the first label or the middle), and   
    - `x_layout = "right"` right-justifies labels (on the first label or the middle).   

2. `y_layout` accepts: 

    - `y_layout = "even"` for an even distribution of labels across the available y-space, and    
    - `y_layout = "flush"` to omit space between adjacent labels (as in `y_layout = 0`).   


Thus, the position and orientation of the text labels can be controlled by using the `x_layout` and `y_layout` options: 

<!-- Example: mark x_layout -->

```{r mark-x-layout-center, echo = FALSE, eval = FALSE, fig.width = 4, fig.asp = 1}
# Labels to plot: 
s <- c("Please note:",
       "These text labels",
       "are centered:", 
       ">>> <<<")

# Create a new plot (of type "slide"):
mark(labels = s, 
     x = NA, y = .85, 
     x_layout = "center", y_layout = "even",
     col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = 1.2,
     new_plot = "slide")

# mark middle:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset
```

<!-- Fine tuning: mark x_layout -->

```{r mark-x-layout-all, echo = FALSE, eval = TRUE, fig.width = 6, fig.height = 3.5, out.width = "60%"}
# (A) x_layout = "center": ---- 
s <- c("Please note:", "These text labels",
       "are centered:", ">>> <<<")

# Create a new plot (of type "slide"):
mark(labels = s, x = NA, y = .85, 
     x_layout = "center", y_layout = "even",
     col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = 1,
     new_plot = "slide")

# mark middle:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset

# (B) x_layout = "right": ----  
s <- c("Please note:", "These text labels",
       "are right-justified:", ">>>")

x_r <- 4/3 - .05

# Add to plot:
mark(labels = s, x = x_r, y = .85, 
     x_layout = "right", y_layout = "even",
     col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = 1)

abline(v = x_r, col = "grey", lty = 2)

# (C) x_layout = "left": ----  
s <- c("Please note:", "These text labels",
       "are left-justified:", "<<<")

x_l <- .05

# Add to plot:
mark(labels = s, x = x_l, y = .85, 
     x_layout = "left", y_layout = "even",
     col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = 1)

abline(v = x_l, col = "grey", lty = 2)
```

See `?mark` for details. 

### Underline text

The `uline()` function allows emphasizing text by plotting it with colored underlining (to provide the functionality of "Unterstreichen"): 

```{r uline-demo, echo = FALSE, fig.width = 3, fig.height = 3, out.width = "35%"}
slogan <- c("Geradlinig", "Authentisch", "Beweglich", "Offen", "Paradiesisch")
uline(labels = slogan,
      x = 0, y = .9, y_layout = "even",
      cex = 1.1, font = 2,
      new_plot = "blank")
```

As with `mark()` above, the `uline()` command distinguishes between creating a new plot (by setting `new_plot` to either `"blank"` or "`slide`") or 
adding to a pre-existing plot (by using the default `new_plot = "none"`): 

```{r uline-ex1, echo = TRUE, fig.width = 4.5, fig.height = 3, out.width = "45%"}
lbl_uline <- c("This is neat, true, and terribly important")
uline(labels = lbl_uline, new_plot = "blank", y = .6, cex = 1.1)  # create a new plot
uline(labels = "(which is why we underline it).", y = .45, cex = 1)  # add to plot
```

The arguments for text elements (`labels`), text positions (`x`, `y`, `y_layout`), and text properties and decorations (`col`, `col_bg`, `cex`, and `font`) work as they do for `mark()`. For instance, shorter vectors are recycled to match the length of `labels`: 

```{r uline-ex2, echo = TRUE, fig.width = 3, fig.height = 3, out.width = "35%"}
slogan <- c("Geradlinig", "Farbenfroh", "Beweglich", "Bunt", "Paradiesisch")
uline(labels = slogan,
      x = c(0, .5), y = .9, y_layout = "even",
      col = c("black", Bordeaux), col_bg = c(Seeblau, Seegruen), 
      cex = 1.1, font = 2, new_plot = "blank")
```

The position and orientation of the text labels can be controlled by using the `x_layout` and `y_layout` options: 

<!-- Example: uline x_layout -->

```{r uline-x-layout-center, echo = FALSE, eval = FALSE, fig.width = 4, fig.asp=1}
# Labels to plot: 
s <- c("Please note:",
       "These text labels",
       "are centered:", 
       ">>> <<<")

# Create a new plot (of type "slide"):
uline(labels = s, 
      x = NA, y = .85, 
      x_layout = "center", y_layout = "even",
      col = "black", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
      cex = 1.0,
      new_plot = "slide")

# mark middle:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset
```

<!-- Fine tuning: uline x_layout -->

```{r uline-x-layout-all, echo = FALSE, eval = FALSE, fig.width = 5, fig.height = 3.5, out.width = "60%"}
# (A) x_layout = "center": ---- 
s <- c("Please note:", "These text labels",
       "are centered:", ">>> <<<")

# Create a new plot (of type "slide"):
uline(labels = s, x = NA, y = .85, 
     x_layout = "center", y_layout = "even",
     col = "black", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = .9, new_plot = "slide")

# mark middle:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset

# (B) x_layout = "right": ----  
s <- c("Please note:", "These text labels",
       "are right-justified:", ">>>")

x_r <- 4/3 - .10

# Add to plot:
uline(labels = s, x = x_r, y = .85, 
     x_layout = "right", y_layout = "even",
     col = "black", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = .9)

abline(v = x_r, col = "grey", lty = 2)

# (C) x_layout = "left": ----  
s <- c("Please note:", "These text labels",
       "are left-justified:", "<<<")

x_l <- .10

# Add to plot:
uline(labels = s, x = x_l, y = .85, 
     x_layout = "left", y_layout = "even",
     col = "black", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
     cex = .9)

abline(v = x_l, col = "grey", lty = 2)
```


### Post text

The `post()` function allows adding text to a rectangular\ `xbox` (to provide the functionality of "Merken"): 

```{r post-demo, echo = FALSE, fig.width = 4, fig.height = 2, out.width = "45%"}
xbox(col = Seegruen, dim = c(4, 2))
post(labels = "Please note", x = .1, y = 1.2, cex = 1.2, font = 2, new_plot = FALSE)
post(labels = c("For best results, always design and view", "boxes with text at the size at which they", "are saved and displayed later."), 
     x = .1, y = .8, y_layout = .1, cex = 1, font = 1, new_plot = FALSE)
```

By default, `post()` assumes a pre-existing `xbox` to which the text elements of `labels` are to be added.  
For square boxes, setting `new_plot` to `"xbox"` creates a new\ `xbox` with its default dimensions of `dim = c(1, 1)`.  
By adjusting its color and text arguments, `post()` allows creating simple badges or logos: 

```{r post-logo, echo = TRUE, fig.width = 2, fig.height = 2, fig.show = "hold", out.width = "25%"}
# Badge (to view as 250 x 250 pixel):
post(labels = "ToDo", y = .35, 
     col = "white", col_bg = pal_signal[[1]],
     cex = 3, font = 2, new_plot = "xbox")  

# Badge (to view as 250 x 250 pixel):
post(labels = "R", y = .35, 
     col = "white", col_bg = pal_seeblau[[5]],
     cex = 8, font = 2, new_plot = "xbox")
```

<!-- Creating course logos: -->

```{r post-logo-hide, echo = FALSE, eval = FALSE, out.width = "25%"}
# To be viewed/saved at 250 x 250 pixels:

# ds4psy course logo: 
post(labels = "ds4psy", x = .05, y = .34, 
     col = "white", col_bg = pal_seeblau[[5]],
     cex = 2.2, font = 2, new_plot = "xbox")

# i2ds course logo: 
post(labels = "i2ds", x = .05, y = .34, 
     col = "white", col_bg = pal_pinky[[4]],
     cex = 2.2, font = 2, new_plot = "xbox")
```

Another potential use of `post()` is for creating images providing contact details: 

```{r post-address, echo = TRUE, fig.width = 3, fig.height = 3, out.width = "35%"}
address <- c("Dr. B. F. Skinner", " ",
             "Department of Psychology",
             "Office Z101",
             "Tel.: +49 7531 88-0815",
             "Fax: +49 7531 88-0810",
             "b.f.skin@uni-konstanz.de")

post(labels = "Contact",
     y = .7,
     cex = 1.2, font = 2,
     col_bg = Petrol,
     new_plot = "xbox")  # create new plot 

post(labels = address,
     y = .5, y_layout = .03,
     new_plot = "none")  # add to plot
```

As above, the position and orientation of the text labels can be controlled by using the `x_layout` and `y_layout` options: 

<!-- Example: post x_layout -->

```{r post-x-layout-center, echo = FALSE, eval = FALSE, fig.width = 3, fig.height = 3, out.width = "35%"}
# Labels to plot: 
s <- c("Please note:",
       "These text labels",
       "are centered:", 
       ">>> <<<")

# Create a new plot (of type "slide"):
post(labels = s, 
     x = NA, y = .80, 
     x_layout = "center", y_layout = "even",
     col = "white", col_bg = c(Karpfenblau),
     font = 2, cex = 1.0,
     new_plot = "xbox")

# mark middle:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset
```

<!-- Fine tuning: post x_layout -->

```{r post-x-layout-all, echo = FALSE, eval = TRUE, fig.width = 3, fig.height = 3, out.width = "35%"}
# (A) x_layout = "center": ---- 
s <- c("Note:", "Labels",
       "centered:", ">>> <<<")

# Create a new plot (of type "slide"):
post(labels = s, x = NA, y = .80, 
     x_layout = "center", y_layout = "even",
     col = "white", col_bg = Karpfenblau,
     cex = .85, new_plot = "xbox")

# middle of current plot:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset


# (B) x_layout = "right": ----  
s <- c("Adjust", "on", "the right:", ">>>")

x_r <- 1 - .15

# Add to plot:
post(labels = s, x = x_r, y = .80, 
     x_layout = "right", y_layout = "even",
     col = "white", 
     cex = .85)

abline(v = x_r, col = "grey", lty = 2)

# (C) x_layout = "left": ----  
s <- c("Adjust", "on", "the left:", "<<<")

x_l <- .15

# Add to plot:
post(labels = s, x = x_l, y = .80, 
     x_layout = "left", y_layout = "even",
     col = "white", 
     cex = .85)

abline(v = x_l, col = "grey", lty = 2)
```



### Headings

The `heading()` function makes it easy to create recognizable headings: 

```{r heading-demo, echo = FALSE, fig.width = 3, fig.height = 2, out.width = "40%"}
heading(labels = c("Ich bin", "eine", "Headline."))
```

By default, the function creates a new slide (i.e., `new_plot = "slide"`) and positions the text elements of\ `labels` on its left edge (i.e., `x = 0`, `y = .8` and `y_layout = "flush"`). 
When headings violate the official recommendations, a message is provided:

```{r heading-warn-1, echo = TRUE, eval = TRUE, fig.align = 'center', fig.width = 3, fig.height = 3, out.width = "40%"}
heading(labels = c("Ewig", "währt am", "längsten..."), 
        y = 2/3, col_bg = usecol(pal_pinky))
```

Re-arranging words helps to avoid step-wise titles (see p.\ 25 of the CD\ manual): 

```{r heading-warn-2, echo = TRUE, message = TRUE, fig.width = 3, fig.height = 2, out.width = "40%"}
# Beware of step-wise arrangements:
heading(labels = c("Ich", "bin keine", "gute Headline."))

# Avoiding warning:
heading(labels = c("Ich bin", "eine alternative", "Headline."))
```

When colors are left unspecified, `heading()` selects colors that are appropriate 
for the number of elements in `labels` (i.e., shades of `Seeblau` for `col_bg = "default"`). 
However, default colors can be overwritten by setting the\ `col` and\ `col_bg` arguments:

```{r heading-col, echo = TRUE, fig.width = 3.5, fig.height = 2.5, out.width = "40%"}
heading(labels = c("Ene,", "mene, miste,", "es rappelt", "in der Kiste."), 
        col = "white", col_bg = c(Pinky, Seegruen, Bordeaux, Karpfenblau))
```

<!-- The position and orientation of the text labels can be controlled by using the `x_layout` and `y_layout` options:  -->

<!-- Example: heading x_layout -->

```{r heading-x-layout-center, echo = FALSE, eval = FALSE, fig.width = 4, fig.asp=1, out.width = "45%"}
# Labels to plot: 
s <- c("Please note:",
       "These headings",
       "are centered:", 
       ">>> <<<")

# Create a new plot (of type "slide"):
heading(labels = s, 
        x = NA, y = .85, 
        x_layout = "center", y_layout = "even",
        col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
        font = 2, cex = 1.2,
        new_plot = "slide")

# mark middle:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset
```

<!-- Fine tuning: heading x_layout -->

```{r heading-x-layout-all, echo = FALSE, eval = FALSE, fig.width = 6, fig.height = 4, out.width = "45%"}
# (A) x_layout = "center": ---- 
s <- c("Please note:", "These headings",
       "are centered:", ">>> <<<")

# Create a new plot (of type "slide"):
heading(labels = s, x = NA, y = .85, 
        x_layout = "center", y_layout = "even",
        col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
        cex = 1,
        new_plot = "slide")

# middle of current plot:
plot_dim <- par("usr")
x_middle <- (plot_dim[2]-plot_dim[1])/2
abline(v = x_middle - 32/600, col = "grey", lty = 2)  # corrected for offset

# (B) x_layout = "right": ----  
s <- c("Please note:", "These headings",
       "are right-justified:", ">>>")

x_r <- 1.23

# Add to plot:
heading(labels = s, x = x_r, y = .85, 
        x_layout = "right", y_layout = "even",
        col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
        cex = 1, new_plot = FALSE)

abline(v = x_r, col = "grey", lty = 2)

# (C) x_layout = "left": ----  
s <- c("Please note:", "These headings",
       "are left-justified:", "<<<")

x_l <- .10

# Add to plot:
heading(labels = s, x = x_l, y = .85, 
        x_layout = "left", y_layout = "even",
        col = "white", col_bg = c(Pinky, Petrol, Bordeaux, Karpfenblau),
        cex = 1, new_plot = FALSE)

abline(v = x_l, col = "grey", lty = 2)
```



### URLs

One of the more obscure rules (on p. 31 of the CD manual) concerns the rendering of URLs, in which typical web patterns (like "http" or "www") are to be omitted and replaced by a typographical dash (which is longer than a hyphen). The `url_unikn` utility function allows formatting URLs the uni.kn way:

```{r url-post, fig.align = 'center', fig.width = 2, fig.height = 2, out.width = "25%"}
my_url <- url_unikn("https://www.uni-konstanz.de/")  # input URL
my_url # formatted URL

# Put URL at bottom of an xbox:
post(labels = my_url, x = .05, y = .1, font = 4, 
     col_bg = pal_seeblau[[5]], new_plot = "xbox")
```


## Resources

The following versions of **unikn** and corresponding resources are currently available: 

Type:                    | Version:           | URL:                           |         
:------------------------|:-------------------|:-------------------------------| 
A. **unikn** (R package): | [Release version](https://CRAN.R-project.org/package=unikn) | <https://CRAN.R-project.org/package=unikn> |
    &nbsp;                | [Development version](https://github.com/hneth/unikn/)       | <https://github.com/hneth/unikn/> | 
B. Online documentation:  | [Release version](https://hneth.github.io/unikn/)            | <https://hneth.github.io/unikn/> | 
    &nbsp;                | [Development version](https://hneth.github.io/unikn/dev/)    | <https://hneth.github.io/unikn/dev/> | 


## Vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://CRAN.R-project.org/package=unikn">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "155px" style = "width: 155px; float: right; border:0;"/>
</a> 

The following vignettes provide instructions and examples for using the **unikn** colors, color palettes, and functions: 

<!-- Table of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color functions |
| 2. | [Color recipes](color_recipes.html) | Recipes for color-related tasks |
| 3. | [Institutional colors](color_inst.html) | Creating color palettes for other institutions |
| 4. | [Text](text.html)     | Text boxes and decorations | 

<!-- eof. -->