---
title: "Text boxes and decorations"
author: "Hansjörg Neth, SPDS, uni.kn"
date: "2019 03 05"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Text}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Beyond providing support for [colors](colors.html), `unikn` provides functions for plotting graphical elements (like boxes) and text decorations (with colored backgrounds or underlining). 
Before we illustrate these functions through examples, it's important to note some caveats: 

- Plotting text (i.e., graphically rendering characters) is rarely a good idea. 
Plotted text typically doesn't scale (when changing the size of images) and cannot be recognized automatically (e.g., copied, indexed, or scraped). 
Hence, the following functions should only be used in contexts in which no better solutions are available 
(e.g., when specifically creating images, or when having to add text to graphs). 

- Like all other templates, our renderings are subject to constraints and limitations. 
As a standard installation of R lacks the prescribed "Theinhardt" fonts, we can only mimic the official design specifications (in Arial, sans serif). 

That said, it's sometimes useful to create digital fluff. Provided that decorative elements can be produced easily and at low cost, they can help to structure and emphasize information. 
The `unikn` package helps preventing common mistakes by novices (e.g., boxes or lines extending beyond text, or step-functions in multi-line titles) and can be customized and improved by expert users. Hence, we hope that the following functions are useful for plotting graphical elements (e.g., boxes, logos, etc.) and help you achieve a uniform look when styling your presentations and visualizations. 

Please load the `unikn` package to get started: 

```{r load_pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```

## Plotting empty slides and boxes

As the text functions of `unikn` add styled text to an existing plot, we need to create a plot to which we can add text. 
`unikn` provides 2 functions to begin a new plot:

1. `slide` plots an empty canvass or frame:

```{r slide_def, fig.align = 'center', fig.width = 4, fig.height = 3}
slide()  # plot default slide
```

2. `xbox` plots a colored box with an X in its top right corner:

```{r xbox_def, fig.align = 'center', fig.width = 2, fig.height = 2}
xbox()  # plot default xbox
```

### Colors

The background and border colors of slides can be changed by setting their `col`, `border`, and `lwd` arguments:

```{r slide_col, fig.align = 'center', fig.width = 4, fig.height = 3}
slide(col = pal_seeblau[[1]], border = Petrol, lwd = 5)  # set colors of slide and border
slide(col = pal_peach[[2]], border = NA)  # colored slide without border
```

Similarly, the background color of xboxes can be changed by the `col` argument:

```{r xbox_col, fig.align = 'center', fig.width = 2, fig.height = 2}
xbox(col = Pinky)  # set color of xbox
```

### Dimensions

By default, both `slide` and `xbox` fill the entire plotting region (i.e., all margins are set to 0).

The width of a slide is set to 4/3 of its height (of 1). 
To change this default, use the `dim` argument for specifying the slide's x- and y-dimensions, respectively:

```{r slide_dim, fig.align = 'center', fig.width = 6, fig.height = 2}
slide(dim = c(3, 1))  # a slide 3 times wider than high
```

Similarly, an xbox has the default dimensions of a unit square (i.e., `dim = c(1, 1)`). 
To change this default, use the `dim` argument for specifying the xbox's x- and y-dimensions, respectively:

```{r xbox_dim, fig.align = 'center', fig.width = 2, fig.height = 4}
xbox(col = Petrol, dim  = c(1, 2))  # an xbox 2 times higher than wide
```

## Plotting styled text

Three main functions:

- `mark`
- `line`
- `post`

In addition, two helper functions: 

- `headline`
- `url_unikn`

### Mark text

"Markieren"

### Underline text

"Unterstreichen"

### Post it

"Merken"

### Headlines

Titels

### URLs

One of the more obscure rules (in p. 31 of the CD manual) concerns the rendering of URLs, in which typical web patterns (like "http" or "www") are to be omitted and replaced by a typographical dash (which is longer than a hyphen). The `url_unikn` utility function allows formatting URLs the uni.kn way:

```{r url_post, fig.align = 'center', fig.width = 2, fig.height = 2}
my_url <- url_unikn("https://www.uni-konstanz.de/")  # input URL as copied from web browser
my_url  # formatted URL

post(lbls = my_url, y = .1, font = 4, new_plot = "xbox")  # post at bottom of an xbox
```


## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color palettes | 
| 2. | [Text](text.html)     | Text boxes and decorations | 


<!-- eof. -->
