---
title: "Text boxes and decorations"
author: "SPDS, uni.kn"
date: "2019 03 24"
output: 
  rmarkdown::html_vignette: 
    fig_caption: yes
vignette: > 
  %\VignetteIndexEntry{Text}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
options(max.print = "75")
knitr::opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               collapse = TRUE, 
               comment = "#>",
               message = FALSE,
               warning = FALSE,
               # ## Default figure options:
               # fig.width = 7.2, 
               # fig.asp = .618, # golden ratio
               # out.width = "75%",
               fig.align = "center"
               )
```

Beyond providing support for [colors](colors.html), `unikn` provides functions for plotting graphical elements (like boxes) and text decorations (with colored backgrounds or underlining). 
Before illustrating these functions, please note some caveats: 

- Plotting text (i.e., graphically rendering characters) is rarely a good idea. In contrast to the vector fonts used by DTP programs, plotted text typically does not scale (when changing the size of images) and cannot be recognized automatically (e.g., copied, indexed, or scraped). 
Hence, the following functions should only be used when no better solutions are available (e.g., for adding text to graphs, or for obstructing automatic text recognition). 

- Like all other templates, our renderings are subject to constraints and limitations. For instance, as any standard installation of [R](https://www.r-project.org/) lacks the prescribed "Theinhardt" fonts, we only mimic the official design specifications (in Arial, sans serif). 

That said, creating digital fluff can sometimes be useful. Provided that decorative elements can be produced easily and flexibly, they can be used to emphasize or structure information. The `unikn` package helps preventing common mistakes by novices (e.g., boxes or lines extending beyond text, or step-functions in multi-line titles) and can be customized and improved by expert users. Hence, we hope that the following functions are useful for plotting graphical elements (e.g., boxes, logos, etc.) and help you achieve a uniform look when styling presentations or visualizations.

Please load the `unikn` package to get started: 

```{r load_pkg, message = FALSE, warning = FALSE}
library(unikn)  # load unikn package
```

## Plotting empty slides and boxes

As the text functions of `unikn` add styled text to an existing plot, we must first create a plot to which we can add text. 
`unikn` provides 2 basic functions for creating new plots:

1. `slide` plots an empty canvass (or rectangular frame):

```{r slide_def, fig.align = 'center', fig.width = 4, fig.height = 3}
slide()  # plot default slide
```

2. `xbox` plots a colored box with an X in its top right corner:

```{r xbox_def, fig.align = 'center', fig.width = 2, fig.height = 2}
xbox()  # plot default xbox
```

Both commands accept additional arguments to customize their appearance. 

### Colors

The background and border colors of slides can be changed by setting their `col`, `border`, and `lwd` arguments:

```{r slide_col, fig.align = 'center', fig.width = 4, fig.height = 3}
slide(col = pal_seeblau[[1]], border = Petrol, lwd = 5)  # set colors of slide background and border
slide(col = pal_peach[[2]], border = NA)  # colored slide without a border
```

Similarly, the background color of xboxes can be changed by the `col` argument:

```{r xbox_col, fig.align = 'center', fig.width = 2, fig.height = 2}
xbox(col = Pinky)  # set background color of xbox
```

### Dimensions

By default, both `slide` and `xbox` fill the entire plotting region (i.e., all margins are set to 0).

The width of a slide is set to 4/3 of its height (of 1). 
To change this default, use the `dim` argument for specifying the slide's x- and y-dimensions, respectively:

```{r slide_dim, fig.align = 'center', fig.width = 6, fig.height = 2}
slide(dim = c(3, 1))  # a slide 3 times wider than high
```

Similarly, an xbox has the default dimensions of a unit square (i.e., `dim = c(1, 1)`). 
To change this default, use the `dim` argument for specifying the xbox's `x`- and `y`-dimensions, respectively:

```{r xbox_dim, fig.align = 'center', fig.width = 2, fig.height = 4}
xbox(col = Petrol, dim  = c(1, 2))  # an xbox 2 times higher than wide
```

The 2 strokes of the X (in the top right corner) are orthogonal when viewing the `xbox` at the aspect ratio indicated by `dim` (i.e., here: at a ratio of width x = 1 to height y = 2). When viewing the box at any different aspect ratio, the X seems distorted. Hence, viewing and printing plots at the desired aspect ratio is crucial for designing an xbox.

## Plotting styled text

There are 3 main functions for plotting text with decorations: 

- `mark` allows highlighting text with colored background boxes; 
- `line` allows underlining text with colored lines; 
- `post` allows posting text in rectangular `xbox` elements.

In addition, there are 2 helper functions: 

- `headline`
- `url_unikn`

The following examples illustrate each function. 

### Mark text

The `mark` function allows highlighting text by plotting it with colored background boxes (to provide the functionality of "Markieren"): 

```{r mark_demo, echo = FALSE, fig.width = 3, fig.height = 3}
lbl_blank <- c("                                                ",
              "                                      ",
              "                                                      ",
              "                                                ",
              "                              ")

# (a) Create a new plot (of type "slide"):
mark(labels = lbl_blank,
     x = 0, y = .85, y_layout = .03,
     col_bg = Seeblau,
     cex = 1.0,
     new_plot = "slide")
```

By default, `mark` assumes that we want to add highlighted text to an already existing plot. 
The text labels to be added are provided (as a character vector) by the `labels` argument, their coordinates as the corresponding `x` and `y` arguments. 
Various text decorations (i.e., text and background colors, text sizes, and font types) can be set by specifying additional arguments (i.e., `col`, `col_bg`, `cex`, and `font`, respectively): 

```{r mark_messy_plot, echo = TRUE, fig.width = 7, fig.height = 5}
# Some pre-existing plot: 
n <- 20
set.seed(1)
plot(x = runif(n), y = runif(n), type = "p", pch = 16, cex = 20, col = grey(0, .2),
     axes = FALSE, xlab = "", ylab = "", mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))

# Add 1 marked label:
mark(labels = "Some messy plot", x = .05, y = .85)

# Add 2 more marked labels:
mark(labels = c("Note something here", "More highlighting here"), 
     x = c(.35, .55), y = c(.15, .40),
     col_bg = c(pal_seeblau[[2]], pal_peach[[3]]), cex = 1.2)
```

If no pre-existing plot is available, setting the `new_plot` argument to either `"blank"` or `"slide"` will generate a new default plot (by using the `slide` function described above):

```{r mark_new_plot, echo = TRUE, fig.width = 4, fig.height = 4}
# A sentence to plot: 
s <- c("Markieren",
       "ist ein Bestandteil",
       "von Studieren.")

# Create a new plot (of type "slide"):
mark(labels = s,
     x = 0, y = .85, y_layout = .03,
     col_bg = Seeblau,
     cex = 1.5,
     new_plot = "slide")

# Add text to the same plot:
mark(labels = s,
     x = 0, y = .40, y_layout = .03,
     col_bg = pal_pinky[[2]],
     cex = 1.5, 
     new_plot = "none")
```

Instead of providing explicit `x` and `y` coordinates for all text elements in `labels`, the calls above defined an initial coordinate and used a numeric `y_layout` argument to fix the distance between consecutive text labels (to a value of `.03`). Special values for `y_layout` are: 

- `y_layout = "even"` for an even distribution of labels across the available y-space, and 
- `y_layout = "flush"` to omit space between adjacent labels (as in `y_layout = 0`).  

See `?mark` for details. 

### Underline text

The `line` function allows emphasizing text by plotting it with colored underlining (to provide the functionality of "Unterstreichen"): 

```{r line_demo, echo = FALSE, fig.width = 3, fig.height = 3}
slogan <- c("Geradlinig", "Authentisch", "Beweglich", "Offen", "Paradiesisch")
line(labels = slogan,
     x = 0, y = .9, y_layout = "even",
     cex = 1.1, font = 2,
     new_plot = "blank")
```

As with `mark` above, the `line` command distinguishes between creating a new plot (by setting `new_plot` to either `"blank"` or "`slide`") or 
adding to a pre-existing plot (by using the default `new_plot = "none"`): 

```{r line_ex1, echo = TRUE, fig.width = 4.5, fig.height = 3}
lbl_line <- c("This is neat, true, and terribly important")
line(labels = lbl_line, new_plot = "blank", y = .6, cex = 1.1)  # create a new plot
line(labels = "(which is why we underline it).", y = .45, cex = 1)  # add to plot
```

The arguments for text elements (`labels`), text positions (`x`, `y`, `y_layout`), and text properties and decorations (`col`, `col_bg`, `cex`, and `font`) work as they do for `mark`. For instance, shorter vectors are recycled to match the length of `labels`: 

```{r line_ex2, echo = FALSE, fig.width = 3, fig.height = 3}
slogan <- c("Geradlinig", "Farbenfroh", "Beweglich", "Bunt", "Paradiesisch")
line(labels = slogan,
     x = c(0, .45), y = .9, y_layout = "even",
     col = c("black", Bordeaux), col_bg = c(Seeblau, Seegruen), 
     cex = 1.1, font = 2,
     new_plot = "blank")
```

### Post it

The `post` function allows adding text to a rectangular `xbox` (to provide the functionality of "Merken"): 

```{r post_demo, echo = FALSE, fig.width = 4, fig.height = 2}
xbox(col = Seegruen, dim = c(4, 2))
post(labels = "Please note", x = .1, y = 1.2, cex = 1.2, font = 2, new_plot = FALSE)
post(labels = c("For good results, always design and view", "boxes and text at the size at which they", "will be saved and displayed later."), 
     x = .1, y = .8, y_layout = .1, cex = 1, font = 1, new_plot = FALSE)
```

By default, `post` assumes a pre-existing `xbox` to which the text elements of `labels` are to be added.  
For square boxes, setting `new_plot` to `"xbox"` creates a new `xbox` with its default dimensions of `dim = c(1, 1)`.  
By adjusting its color and text arguments, `post` allows creating simple badges or logos: 

```{r post_logo, echo = TRUE, fig.width = 2, fig.height = 2, fig.show = "hold"}
# Badge (to view as 250 x 250 pixel):
post(labels = "ToDo", y = .35, 
     col = "white", col_bg = pal_signal[[1]],
     cex = 3, font = 2, new_plot = "xbox")  

# Badge (to view as 250 x 250 pixel):
post(labels = "R", y = .35, 
     col = "white", col_bg = pal_seeblau[[5]],
     cex = 8, font = 2, new_plot = "xbox")
```

Another use of `post` is to create an image that provides contact details: 

```{r post_address, echo = TRUE, fig.width = 3, fig.height = 3}
address <- c("Dr. B. F. Skinner", " ",
             "Department of Psychology",
             "Office F101",
             "Tel.: +49 7531 88-0815",
             "Fax: +49 7531 88-0810",
             "b.f.skin@uni-konstanz.de")

post(labels = "Contact",
     y = .7,
     cex = 1.2, font = 2,
     col_bg = Petrol,
     new_plot = "xbox")  # create new plot 

post(labels = address,
     y = .5, y_layout = .03,
     new_plot = "none")  # add to plot
```

### Headlines

Titels

+++ here now +++


### URLs

One of the more obscure rules (in p. 31 of the CD manual) concerns the rendering of URLs, in which typical web patterns (like "http" or "www") are to be omitted and replaced by a typographical dash (which is longer than a hyphen). The `url_unikn` utility function allows formatting URLs the uni.kn way:

```{r url_post, fig.align = 'center', fig.width = 2, fig.height = 2}
my_url <- url_unikn("https://www.uni-konstanz.de/")  # input URL as copied from web browser
my_url  # formatted URL

post(labels = my_url, y = .1, font = 4, new_plot = "xbox")  # post at bottom of an xbox
```


## All unikn vignettes

<!-- unikn pkg logo and link: -->
<a href = "https://github.com/hneth/unikn/">
<img src = "../inst/pix/unikn.png" alt = "unikn::" align = "right" width = "140px" style = "width: 140px; float: right; border:0;"/>
</a>

<!-- Index of vignettes: -->

| Nr.  | Vignette | Content    |        
| ---: |:---------|:-----------|
| 1. | [Colors](colors.html) | Colors and color palettes | 
| 2. | [Text](text.html)     | Text boxes and decorations | 


<!-- eof. -->
